---
title: "CBO_Seurat_EXPLORE1_50K"
author: "JJ"
date: "07/08/2022"
output: html_document
---

```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE,
                      fig.path = "./outputs50K/CBO/figures/")
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(glmGamPoi)
library(RColorBrewer)
library(Matrix)
library(clustree)
library(ggVennDiagram)
library(patchwork)
library(tidyverse)  
})

```


```
{r}
load("./RDataFiles/CBO_Seurat_EXPLORE1_50K.RData")
```


```
# load full filtered (genes and cells) seurat object

Fullfilt.cbo.seurat <- readRDS("outputs50K/CBO/filtered.cbo.seurat.rds")
```

```
{r}
# load the single CBO control sample (CCdifference regressed out!), already SCTransformed,
# then run PCA, UMAP, etc on 'phase.filtRiboMito.ons76cbo.SCT.regressedCCdiff.rds'
# which is saved in the outputs50K folder and can be accessed by readRDS above. 

phase.cbo.sct <- readRDS("outputs50K/CBO/phase.cbo.filtRiboMito.SCT.CCdiffRegress.rds")

# run PCA, UMAP etc on 'phase.cbo.sct' object from the CBO sample
phase.cbo.sct <- phase.cbo.sct %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE)
```

```
{r DimPlotPhaseCBOsctCCdiffRegressClustered}
DimPlot(phase.cbo.sct, label = TRUE)
```

```
# save the SCTransformed clustered phase CBO object
saveRDS(phase.cbo.sct, 
        file = "outputs50K/CBO/phase.filtRiboMito.cbo.SCT.CCdiffRegressed.clustered.rds")
```

```{r}
# load the SCTransformed, fully filtered seurat object
phase.cbo.sct.clusters <- 
  readRDS("outputs50K/CBO/phase.filtRiboMito.cbo.SCT.CCdiffRegressed.clustered.rds")
```

```{r ElbowPlotGeneCellFiltCBO}
# elbow plot shows that there is an elbow at PC 15, but the graph does not level off after
# PC 15 showing there is some additional information to ~ PC 40.
# The input object is the CLUSTERED saved seurat object

ElbowPlot(phase.cbo.sct.clusters, ndims = 50, reduction = "pca")
```

```
# DO NOT RUN THESE CODE CHUNKS

# add percent.ribo to the metadata (not informative!)
filt.cbo.sct.clusters[["percent.ribo"]] <- PercentageFeatureSet(filt.cbo.sct.clusters, pattern = "'^RP[SL]'")
```

```
# plot metadata features - percent mitochondria (< 10%) and percent ribosome (= 0)
VlnPlot(filt.cbo.sct.clusters, features = c("percent.mt", "percent.ribo"), ncol = 2)
```

----------------------------------------------------------------------------------------------
***Get the differentially expressed genes in the 12 ONS76-CBO clusters using a FOR loop***
```
# DO NOT RUN THESE CODE CHUNKS

Idents(phase.cbo.sct.clusters) <- "seurat_clusters"
markers.pos <- data.frame()
for (i in levels(Idents(phase.cbo.sct.clusters))){
  markers <- FindMarkers(phase.cbo.sct.clusters, ident.1 = i, grouping.var = "ident", 
                         verbose = FALSE, only.pos = TRUE, logfc.threshold = 0.25)  # base 2 log fold change
  dat <- markers %>% rownames_to_column(var = "genes")
  dat$cluster_ID <- i
  dat$cluster_ID <- as.numeric(dat$cluster_ID)
  df <- data.frame(dat)
  markers.pos <- rbind(markers.pos, df)
}
head(markers.pos)
```

```{r}
# Cell counts per cluster.
# there are 3070 cells in total, which is correct, and the cell numbers per cluster are shown below.
# Two clusters are present with < 100 cells

cell.num <- table(phase.cbo.sct.clusters@active.ident)
cell.num
sum(cell.num)
```

```
# DO NOT RUN THESE CODE CHUNKS
dim(markers.pos)
```



```
# save the postive markers of CBO clusters generated from SCTransformed, CC difference,
# ribosomal and mitochondrial gene filtered object:
# 'phase.filtRiboMito.cbo.SCT.CCdiffRegressed.clustered.rds'

saveRDS(markers.pos, file = "outputs50K/CBO/pos.markers.phase.cbo.filtRiboMito.CCdiffRegress.SCT.rds")
```
```{r}
markers.pos <- readRDS("outputs50K/CBO/pos.markers.phase.cbo.filtRiboMito.CCdiffRegress.SCT.rds")
```

Check out the first few DEGs by group (cluster_ID) of 'markers.pos'
```{r}
markers.pos %>% 
  group_by(cluster_ID) %>% 
  slice_head(n = 40)
```

```{r}
markers.pos %>% 
  filter(genes %in% c("KIRREL3"))
```

```{r}
markers.pos %>%
  filter(cluster_ID == "4")
```


```{r}
top10 <- markers.pos %>% 
  group_by(cluster_ID) %>% 
  top_n(10, -p_val_adj)
```

```{r}
top10
```

```{r}
# replace the cluster numbers with the identity of each cluster
levels(phase.cbo.sct.clusters)
```

```{r DimplotCBOcellTypesLabelled, fig.height=6}
# continue to replace the cluster numbers with the identity of each cluster
new.cluster.ids <- c("RP-choroid plexus", "choroid plexus", "glia", "RL/DCN",
                     "uncertain", "granule neuron", "glio-neuronal", "granule progenitor", "RL-derived",
                     "oligodendrocyte", "purkinje neuron", "ependyma")
names(new.cluster.ids) <- levels(phase.cbo.sct.clusters)
phase.cbo.sct.clusters.2 <- RenameIdents(phase.cbo.sct.clusters, new.cluster.ids)
DimPlot(phase.cbo.sct.clusters.2, reduction = "umap", label = TRUE, repel = TRUE, pt.size = 0.5)
```

```{r DimPlotPhaseCBOsctClusters, fig.height=6}
# there are 12 clusters (0 - 11)
DimPlot(phase.cbo.sct.clusters, label = TRUE, repel = TRUE)
```

```{r}
table(phase.cbo.sct.clusters.2$seurat_clusters)
```


```{r DotplotSeppTFmarkers}
# using the Sepp et al paper (Extended Fig 8g), test if markers from that paper are expressed 
# in the seurat object filtered to remove all ribosomal and mitochondrial genes, and where cells which
# have mitoch genes > 10% are excluded.
# MUST use the already clustered seurat object.
# I used just the transcription factor (TF) markers (sepp.TF.markers)

sepp.TF.markers <- c("PAX3", "RFX4", "TCFL71", "GLIS3", "SOX6", "TCF7L2", "HOPX", "TSC22D4", "ID4",
             "ESRRB", "FOXP2", "FOXP4", "MEF2C",
             "PAX2", "TFAP2A", "TFAP2B", "BHLHE22", "PRDM8",
             "PAX6", "BARHL1", "E2F3", "ST18", "SATB2", "ZBTB18", "NEUROD1", "ETV1", "KLF9", "NR3C1")


DotPlot(object = phase.cbo.sct.clusters, features = sepp.TF.markers) +
  coord_flip()
```

```{r DotplotSeppMarkersAndCiliaManyCBO, fig.height=10}
# from the Sepp paper repeat Dotplot with TF and some key non-TF markers
# added markers of ciliated cells from cluster10 above.
# used the argument col.min = 0 so that average expression scale is from 0 to 2.5 in key

sepp.all.markers <- c("TFAP2A", "PAX3", "RFX4", "TCFL71", "TCF7L2", "GLIS3", "SOX6", "HOPX", "TSC22D4",  
                      "ID4", "NCKAP5", "SLIT2", "CLYBL", "SLC1A3", "SLC6A11", "AQP4",  #astroglial/progenitor
                      
                      "ESRRB", "FOXP1", "FOXP2", "FOXP4", "MEF2C", "RORB", "CDH9", "ETV1", "EBF1", "EBF2", 
                      "SKOR2", "DAB1", "IPTR",   # Purkinje
                      
                      "NOTCH1",    # VZ neuroblast
             
                      "PAX2", "NFIA", "NFIB", "TFAP2B", "BHLHE22", "PRDM8",  # interneuron
                      
                      "ZFHX3", "ZFHX4",   # GABAergic deep nuclei neurons
             
                      "TOP2A", "ATOH1", "LMX1A", "OTX2","ETV6", "PAX6", "BARHL1", "E2F3", "ST18", "SATB2", 
                      "ZBTB18", "NEUROD1", "ZIC1", "ZIC4", "GABAR6", "CNTN1", "CNTN2", "KLF9", "NR3C1", "LMO3", 
                      "SYNPR", "KCNIP4", "HCRTR2", 
                      "EOMES", "TRPC3", "GRM1")
                      # UBC/GCP/granule/glut deep cerebellar nuclei                                                                   
seppAll.cilia.stem.markers <- c(sepp.all.markers, "CCNO", "CFAP299", "ROPN1L", "C11orf88", "ANKRD66", "C20orf85", 
                                "STOML3", "TEKT4", "SNTN", "MEIS1", "POU3F2", "HOXB2", "MEIS2", "GRIA4")

DotPlot(object = phase.cbo.sct.clusters, features = seppAll.cilia.stem.markers, col.min = 0) +
  coord_flip()
```

```{r ViolinPlotGranuleLineageByClusterCBO, fig.height=10}
# plot of the clusters containing granule/RL lineage suggests RL-derived clusters are 6, 7, 8
# CDH5 (marker of ONS76 cells) is not expressed but CDH1 is (also a marker of ONS76 cells).
# CD44 is not expressed or expressed very little.
# ITGB1 is expressed at quite a high level across most cell types
# In CBO there is only 1 NEUROD1, RBFOX3, RTN1 cluster

VlnPlot(phase.cbo.sct.clusters, features = c("NEUROD1", "CNTN1", "CNTN2", "RBFOX3", "RTN1", "ZIC1"))
```


```{r VlnPlotStemGenesPhaseCBOsct, fig.height=24, fig.width=12}
VlnPlot(phase.cbo.sct.clusters, features = c("HOXB2", "IRX3", "GRIA4", "IRX5", "TWIST1", "PAX3", 
                                             "MIR124-2HG", "PTCH1", "GLI2",
                                             "POU3F2", "MEIS1", "LHX9", "CHCHD2", "PRRX1", "DLL1", "DLL3",
                                             "DLL4","HES5", "HES6", "JAG1", "JAG2", "NNAT", "TTYH1", "CD24"))
```

```{r NrgErbb4PhaseCBOsct, fig.height=8, fig.width=12}
VlnPlot(phase.cbo.sct.clusters, features = c("NRG1", "NRG2", "NRG3", "NRG4", "ERBB4"))
```


***Clustree***
```{r}
# attempt to use Clustree package to find the right number of clusters
# -----------------------------------------------------------------------------------------------

# Select a range of resolutions
resolution.range <- seq(from = 0, to = 1.6, by = 0.2)

# Find clusters using a range of resolutions, the clustering information is in the metadata slot
cbo.clusters <- FindClusters(object = phase.cbo.sct.clusters, 
                                      resolution = resolution.range,
                                      verbose = FALSE)
```

```{r ClustreeCBO, fig.height=12}
# the clustering tree finds 9 clusters before some nodes have multiple incoming edges.
# note that at res = 1 there are 14 clusters

clustree(cbo.clusters)
```

-----------------------------------------------------------------------------------------
***Plot the distribution of proliferative cells in the DimPlot***
Superimpose cycling cells onto the UMAP plot of cell clusters

```{r}
phase.cbo.sct <- 
  readRDS("outputs50K/CBO/phase.filtRiboMito.cbo.SCT.CCdiffRegressed.clustered.rds")
```

Get the list of genes expressed in cycling cells (Seurat list from Tirosh et al.). This shows that there are much smaller numbers of actively cycling cells than I first assumed.
```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
prolif.genes <- list(c(s.genes, g2m.genes))
```


```{r}
phase.cbo.sct <- AddModuleScore(phase.cbo.sct,
                        features = prolif.genes,
                        name = "prolif.genes")
FeaturePlot(phase.cbo.sct, features = 'prolif.genes1', label = TRUE, repel = TRUE) +
   scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))
```

Be careful with the colnames of meta.data. Even though I named "prolif.genes" above, getting the colnames has shown that this becomes "prolif.genes1" !!!
```{r}
names(x = phase.cbo.sct[[]])
```

Subset the object to extract cluster 5 cells, which are Neurod1+.
```{r}
neurod1.cbo <- subset(phase.cbo.sct, idents = c("5"))
```

```{r}
dim(neurod1.cbo)
```

```
# save the neurod1 cluster 5 (which was generated from phase cell cycle scored, CC difference regressed,
# SCTransformed seurat object)
saveRDS(neurod1.cbo, file = "./outputs50K/CBO/neurod1Clust.phaseCCDiffRegress.SCT.cbo.rds")
```

```
save.image("./RDataFiles/CBO_Seurat_EXPLORE1_50K.RData")
```


--------------------------------------------------------------------------------------------
***Repeat clustering of CBO dataset using conventional normalization***
```
# NO NEED TO RUN THESE CODE CHUNKS

# load the dataset
filt.RiboMito.cbo <- readRDS("outputs50K/CBO/filt.RiboMitoCellsGenes.seurat.cbo.rds")

# load cell cycle genes in advance of regressing out CC difference
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

filt.RiboMito.cbo <- NormalizeData(filt.RiboMito.cbo, normalization.method = "LogNormalize", scale.factor = 10000)
filt.RiboMito.cbo <- FindVariableFeatures(filt.RiboMito.cbo, selection.method = "vst", nfeatures = 2000)

# regress CC difference
filt.RiboMito.cbo <- CellCycleScoring(phase.cbo.sct,
                g2m.features = g2m.genes,
                s.features = s.genes)
filt.RiboMito.cbo$CC.difference <- filt.RiboMito.cbo$S.Score - filt.RiboMito.cbo$G2M.Score

filt.RiboMito.cbo <- ScaleData(filt.RiboMito.cbo, vars.to.regress = "CC.difference")

# run the rest of the seurat workflow
filt.RiboMito.cbo <- RunPCA(filt.RiboMito.cbo, features = VariableFeatures(object = filt.RiboMito.cbo))
filt.RiboMito.cbo <- FindNeighbors(filt.RiboMito.cbo, dims = 1:10)
filt.RiboMito.cbo <- FindClusters(filt.RiboMito.cbo, resolution = 0.8) # this res gives only 1 NeuroD1 cluster

filt.RiboMito.cbo <- RunUMAP(filt.RiboMito.cbo, dims = 1:10)
DimPlot(filt.RiboMito.cbo, reduction = "umap")
```

```
VlnPlot(filt.RiboMito.cbo, features = c("NEUROD1", "RBFOX3", "RTN1"))
```

This code retained just to show arguments to read a csv file
```
# load 'cbo.cluster.markers.NoRiboNoMito.csv'

cbo.cluster.markers.noRiboMito <- read.csv("outputs50K/CBO/cbo.cluster.markers.NoRiboNoMito.csv", 
                                       header = TRUE, 
                                       stringsAsFactors = FALSE)
```

This code is retained just to show the usage of write.csv
```
# save cbo.cluster.markers.ribo20 as a csv file. This object has percent.ribo < 20

write.csv(cbo.cluster.markers.ribo20, "./outputs50K/CBO/cbo.cluster.markers.ribo20.csv", quote = F)
```









