---
title: "CBO_Seurat_EXPLORE1_50K"
author: "JJ"
date: "07/08/2022"
output: html_document
---

```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE,
                      fig.path = "./outputs50K/CBO/figures/")
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(glmGamPoi)
library(RColorBrewer)
library(Matrix)
library(ClusterMap)
library(clustree)
library(patchwork)
library(tidyverse)  
})
```


```
{r}
load("./RDataFiles/CBO_Seurat_EXPLORE1_50K.RData")
```

```{r}
# load the clusterMap cell, gene, ribo, mito, CC.difference regressed, SCTransformed seurat object
cbo <- 
  readRDS("outputs50K/CBO/ClusterMapInput.Cbo.NoRiboNoMito.SCT.clustered.CCdiffRegressed.rds")
```


```{r DimPlotCBOsample, fig.height=6, fig.width=10}
DimPlot(cbo, label = TRUE)
```


```{r ElbowPlotGeneCellFiltCBO}
# elbow plot shows that there is an elbow at PC 10
# The input object is the CLUSTERED saved seurat object

ElbowPlot(cbo, ndims = 50, reduction = "pca")
```

```{r}
# Cell counts per cluster.
# there are 3049 cells in total, which is correct, and the cell numbers per cluster are shown below.
# Two clusters, 9 and 10 are present with < 100 cells

cell.num <- table(cbo@active.ident)
cell.num
sum(cell.num)
```

***CBO differential markers***
```{r}
# load the cbo markers
cbo.markers <- read.csv("outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.csv",
                        header = TRUE,
                        stringsAsFactors = FALSE)
```

```{r}
# look at the top markers of each CBO cluster
cbo.markers %>% 
  group_by(cluster) %>% 
  slice_head(n = 10)
```

```{r}
# replace the cluster numbers with the identity of each cluster
levels(cbo)
```

***CBO cluster annotation***
```{r}
# continue to replace the cluster numbers with the identity of each cluster - granule precursors
# are so-called because GSEA suggests these care cycling cells
new.cluster.ids <- c("roof plate", "4th vent/choroid plexus", "Purkinje cells_1", "GABA-ergic neurons",
                     "uncertain", "granule neurons", "Purkinje cells_2", "RL/roof plate", "RL-derived",
                     "oligodendrocytes", "Purkinje cells_3", "ciliated cells", "Purkinje cells_4")
names(new.cluster.ids) <- levels(cbo)
cbo.annotate <- RenameIdents(cbo, new.cluster.ids)
```

```{r}
# Create metadata dataframe
metadata <- cbo.annotate@meta.data
# Add cell IDs to metadata
metadata$cells <- rownames(metadata)
# Create sample column
metadata$celltype <- NA
metadata$celltype[which(metadata$seurat_clusters == "0")] <- "roof plate"
metadata$celltype[which(metadata$seurat_clusters == "1")] <- "4th vent/choroid plexus"
metadata$celltype[which(metadata$seurat_clusters == "2")] <- "Purkinje cells_1"
metadata$celltype[which(metadata$seurat_clusters == "3")] <- "GABA-ergic neurons"
metadata$celltype[which(metadata$seurat_clusters == "4")] <- "uncertain"
metadata$celltype[which(metadata$seurat_clusters == "5")] <- "granule neurons"
metadata$celltype[which(metadata$seurat_clusters == "6")] <- "Purkinje cells_2"
metadata$celltype[which(metadata$seurat_clusters == "7")] <- "RL/roof plate"
metadata$celltype[which(metadata$seurat_clusters == "8")] <- "RL-derived"
metadata$celltype[which(metadata$seurat_clusters == "9")] <- "oligodendrocytes"
metadata$celltype[which(metadata$seurat_clusters == "10")] <- "Purkinje cells_3"
metadata$celltype[which(metadata$seurat_clusters == "11")] <- "ciliated cells"
metadata$celltype[which(metadata$seurat_clusters == "12")] <- "Purkinje cells_4"
```

```{r}
cbo.annotate@meta.data <- metadata
```

```{r}
# save the cbo.annotate object
saveRDS(cbo.annotate, 
        file = "./outputs50K/CBO/cbo.filtRiboMito.CellAnnot.SCT.CCdiffRegressed.clustered.clusterMap.rds")
```

```{r}
# load the 'cbo.annotate' object
cbo.annotate <- 
  readRDS("outputs50K/CBO/cbo.filtRiboMito.CellAnnot.SCT.CCdiffRegressed.clustered.clusterMap.rds")
```

```{r DimplotCBOcellTypesAnnotated, fig.height=6, fig.width=10}
DimPlot(cbo.annotate, label = TRUE, repel = TRUE)
```

```{r}
table(cbo.annotate$celltype)
```


Granule marker genes are from DOI: https://doi.org/10.7554/eLife.37551 (Heintz lab)
```{r ViolinPlotGranuleLineageByClusterCBO, fig.height=9, fig.width=12}
# CDH5 (marker of ONS76 cells) is not expressed but CDH1 is (also a marker of ONS76 cells).
# CD44 is not expressed or expressed very little.
# ITGB1 is expressed at quite a high level across most cell types
# In CBO there is only 1 NEUROD1, RBFOX3, RTN1 cluster
# CALB2, SLC17A7, RELN, GABRA6, AQP7 are all granule markers that are not expressed in cluster 5 (NEUROD1+)

VlnPlot(cbo.annotate, 
        features = c("NEUROD1", "RBFOX3", "RTN1", "CDH15"))
```


```{r VlnPlotGria4Twist1PhaseCBOannotate, fig.width=14}
# HOXB2 is not expressed, and none of these markers show up as a violin

VlnPlot(cbo.annotate, 
        features = c("IRX3", "GRIA4", "IRX5", "TWIST1", "PAX3"))
```

```{r VlnPlotGli2Ptch1Pou3F2PhaseCBOannotate, fig.width=14}
VlnPlot(cbo.annotate, features = c("MIR124-2HG", "PTCH1", "GLI2",
                                             "POU3F2", "MEIS1", "LHX9"))
```

```{r VlnPlotDll3Hes5PhaseCBOannotate, fig.width=14}
VlnPlot(cbo.annotate, 
        features = c("PRRX1", "DLL1", "DLL3", "DLL4","HES5"))
```

```{r VlnPlotNotchGenesPhaseCBOannotate, fig.width=14}
VlnPlot(cbo.annotate, 
        features = c("HES6", "JAG1", "JAG2", "NNAT", "TTYH1", "CD24"))
```

```{r VlnPlotNrgErbb4PhaseCBOannotate, fig.height=9, fig.width=12}
VlnPlot(cbo.annotate, features = c("NRG1", "NRG2", "NRG3", "NRG4", "ERBB4"))
```


***Clustree***
```{r}
# attempt to use Clustree package to find the right number of clusters. Not hugely useful!
# -----------------------------------------------------------------------------------------------

# Select a range of resolutions
resolution.range <- seq(from = 0, to = 1.6, by = 0.2)

# Find clusters using a range of resolutions, the clustering information is in the metadata slot
cbo.clusters <- FindClusters(object = cbo, 
                                      resolution = resolution.range,
                                      verbose = FALSE)
```

```{r ClustreeCBO, fig.height=8,fig.width=8}
# the clustering tree finds 9 clusters before some nodes have multiple incoming edges.
# note that at res = 1 there are 14 clusters

clustree(cbo.clusters)
```


-----------------------------------------------------------------------------------------
***ClusterMap to compare the clusters of CBO, DAOY-CBO and ONS76-CBO***

Combine the CBO and DAOY-CBO samples as shown in the example of https://xgaoo.github.io/ClusterMap/pre_analysis/pre_analysis.html

```{r}
# load the CBO raw data (from "working_data" folder)
data_dir <- "./working_data/CBO/"   # the dot is current dir and is essential!
list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r}
# note these are .gz files which can be directly read. n = 3838 cells

cbo.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                         features = paste0(data_dir, "features.tsv.gz"), 
                         cells = paste0(data_dir, "barcodes.tsv.gz"))
```

```{r}
# load the DAOY-CBO raw data
data_dir <- "./working_data/DAOY-CBO/"   # the dot is current dir and is essential!
list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r}
# note these are .gz files which can be directly read. There are 3654 cells and 36,601 genes.

dyCbo.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                      features = paste0(data_dir, "features.tsv.gz"), 
                      cells = paste0(data_dir, "barcodes.tsv.gz"))
```

****ClusterMap input: preparation of the COMBINED CBO and DAOY-CBO object****
```{r}
colnames(cbo.data) = paste0('cbo-', colnames(cbo.data))
colnames(dyCbo.data) = paste0('dyCbo.data-', colnames(dyCbo.data))
```

```{r}
cbo.dyCbo.data = cbind(cbo.data, dyCbo.data)
```


```{r}
cbo.dyCbo.combined <- CreateSeuratObject(counts = cbo.dyCbo.data,
                                         min.cells = 10, min.genes=200,
                                         project = "cbo.dyCbo.comb")

dim(cbo.dyCbo.combined) #24,165 genes and 7492 cells
```

```{r}
# identify mitochondrial genes
mito.genes <- grep(pattern = "^MT", x = rownames(x = cbo.dyCbo.combined), value = TRUE)
percent.mito <- 
  Matrix::colSums(cbo.dyCbo.combined[mito.genes, ])/Matrix::colSums(cbo.dyCbo.combined)

# identify ribosomal genes
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(x = cbo.dyCbo.combined), value = TRUE)
percent.ribo <- 
  Matrix::colSums(cbo.dyCbo.combined[ribo.genes, ])/Matrix::colSums(cbo.dyCbo.combined)

```

```{r}
# add the ribosome and mitochondrial gene percent to the metadata
cbo.dyCbo.combined <- 
  AddMetaData(object = cbo.dyCbo.combined, metadata = percent.mito, col.name = "percent.mito")

cbo.dyCbo.combined <- 
  AddMetaData(object = cbo.dyCbo.combined, metadata = percent.ribo, col.name = "percent.ribo")
```

```{r}
# remove ribosomal and mitochondrial genes
# remove ribosomal genes first
cbo.dyCbo.combined <- cbo.dyCbo.combined[ ! grepl('^RP[SL]', rownames(cbo.dyCbo.combined)), ]

# remove mitochondrial genes 
cbo.dyCbo.combined <- cbo.dyCbo.combined[ ! grepl("^MT-", rownames(cbo.dyCbo.combined)), ]
```

```{r}
dim(cbo.dyCbo.combined) # 24,055 genes and 7492 cells
```

```{r}
# remove all cells with percent.mito > 10%
cbo.dyCbo.combined <- subset(cbo.dyCbo.combined, subset = percent.mito < 0.1)
```

```{r}
dim(cbo.dyCbo.combined) # 24,055 genes and 6086 cells
```

```{r}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference

# this normalisation step recommended prior to CellCycle scoring in Seurat GitHub issue https://github.com/satijalab/seurat/issues/1679
cbo.dyCbo.combined <- NormalizeData(cbo.dyCbo.combined)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
cbo.dyCbo.combined <- CellCycleScoring(cbo.dyCbo.combined,
                g2m.features = g2m.genes,
                s.features = s.genes)
cbo.dyCbo.combined$CC.difference <- cbo.dyCbo.combined$S.Score - cbo.dyCbo.combined$G2M.Score
```

```{r}
cbo.dyCbo.combined <- SCTransform(cbo.dyCbo.combined, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi", 
                   verbose = FALSE)
```

```{r DimPlotClusterMapCombinedObjectDaoyCboAndCbo, fig.height=6, fig.width=10}
cbo.dyCbo.combined <- cbo.dyCbo.combined %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE)
DimPlot(cbo.dyCbo.combined, label = TRUE, repel = TRUE)
```

```{r}
# save the above combined object
saveRDS(cbo.dyCbo.combined, 
        file = "./outputs50K/DAOY-CBO/clusterMap_files/ClusterMapInput.DaoyCbo.Cbo.combined.NoRiboNoMito.SCT.clustered.CCdiffRegressed.rds")
```


****ClusterMap: input preparation of the CBO object****
```{r}
# create the seurat object, keep genes expressed in 10 or more cells.
cbo <- CreateSeuratObject(counts = cbo.data,
                          min.cells = 10, min.genes=200,
                          project = "cbo")

dim(cbo) # 21,582 genes and 3838 cells
```

```{r}
# identify mitochondrial genes
mito.genes <- grep(pattern = "^MT", x = rownames(x = cbo), value = TRUE)
percent.mito <- 
  Matrix::colSums(cbo[mito.genes, ])/Matrix::colSums(cbo)

# identify ribosomal genes
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(x = cbo), value = TRUE)
percent.ribo <- 
  Matrix::colSums(cbo[ribo.genes, ])/Matrix::colSums(cbo)
```

```{r}
# add the ribosome and mitochondrial gene percent to the metadata
cbo <- 
  AddMetaData(object = cbo, metadata = percent.mito, col.name = "percent.mito")

cbo <- 
  AddMetaData(object = cbo, metadata = percent.ribo, col.name = "percent.ribo")
```

```{r}
# remove ribosomal and mitochondrial genes
# remove ribosomal genes first
cbo <- cbo[ ! grepl('^RP[SL]', rownames(cbo)), ]

# remove mitochondrial genes 
cbo <- cbo[ ! grepl("^MT-", rownames(cbo)), ]

dim(cbo) # 21,572 genes and 3838 cells
```

```{r}
# remove all cells with percent.mito > 10%
cbo <- subset(cbo, subset = percent.mito < 0.1)
dim(cbo) # reduced to 21,572 genes and 3048 cells
```

```{r}
# save the NON-normalised, NON-clustered cbo seurat object
saveRDS(cbo, file = "./outputs50K/CBO/filt.RiboMitoCellsGenes.seurat.cbo.rds")
```

```{r}
cbo <- readRDS("outputs50K/CBO/filt.RiboMitoCellsGenes.seurat.cbo.rds")
```


```{r}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference

# normalise object before running cell cycle scores as per Seurat github issue: https://github.com/satijalab/seurat/issues/1679
cbo <- NormalizeData(cbo)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
cbo <- CellCycleScoring(cbo,
                g2m.features = g2m.genes,
                s.features = s.genes)
cbo$CC.difference <- cbo$S.Score - cbo$G2M.Score
```

```{r}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
cbo.input.integration <- SCTransform(cbo, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```


```{r}
# save the SCTransformed object, CC difference regressed, PCA reduction, as input for integrated workflow
saveRDS(cbo.input.integration, "./outputs50K/CBO/cbo.inputIntegration.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```


```{r DimPlotClusterMapCbo, fig.height=6, fig.width=10}
cbo <- SCTransform(cbo, vars.to.regress = "CC.difference", method = "glmGamPoi", verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE)
```

```{r}
# save the clustered, SCTransformed object above
saveRDS(cbo, file = "./outputs50K/CBO/ClusterMapInput.Cbo.NoRiboNoMito.SCT.clustered.CCdiffRegressed.rds")
```

```{r}
# load the above object. This object not yet been annotated.
cbo.clust <- readRDS("outputs50K/CBO/ClusterMapInput.Cbo.NoRiboNoMito.SCT.clustered.CCdiffRegressed.rds")
```

```{r DimplotCboUmapSCTransform, fig.height=6, fig.width=10}
DimPlot(cbo.clust, label = TRUE)
```


```{r}
# get the markers of each cluster
cbo.markers <- FindAllMarkers(object = cbo.clust, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# save the markers of each cluster in CBO as csv file
write.csv(cbo.markers, file = "./outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.csv")
```

```{r}
# save the markers of each cluster in CBO also as RDS file for downstream use
saveRDS(cbo.markers, file = "./outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.rds")
```

```{r}
# load the cbo markers
cbo.markers <- read.csv("outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.csv",
                        header = TRUE,
                        stringsAsFactors = FALSE)
```

```{r}
# look at the top markers of each CBO cluster
cbo.markers %>% 
  group_by(cluster) %>% 
  slice_head(n = 30)
```

```{r}
cbo.markers %>% 
  filter(cluster == "0")
```

****ClusterMap input: preparation of the DAOY-CBO object****

```{r}
dyCbo <- CreateSeuratObject(counts = dyCbo.data,
                          min.cells = 10, min.genes=200,
                          project = "dyCbo")

dim(dyCbo) # 21,928 genes and 3654 cells
```

```{r}
# identify mitochondrial genes
mito.genes <- grep(pattern = "^MT", x = rownames(x = dyCbo), value = TRUE)
percent.mito <- 
  Matrix::colSums(dyCbo[mito.genes, ])/Matrix::colSums(dyCbo)

# identify ribosomal genes
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(x = dyCbo), value = TRUE)
percent.ribo <- 
  Matrix::colSums(dyCbo[ribo.genes, ])/Matrix::colSums(dyCbo)
```

```{r}
# add the ribosome and mitochondrial gene percent to the metadata
dyCbo <- 
  AddMetaData(object = dyCbo, metadata = percent.mito, col.name = "percent.mito")

dyCbo <- 
  AddMetaData(object = dyCbo, metadata = percent.ribo, col.name = "percent.ribo")
```

```{r}
# remove ribosomal and mitochondrial genes
# remove ribosomal genes first
dyCbo <- dyCbo[ ! grepl('^RP[SL]', rownames(dyCbo)), ]

# remove mitochondrial genes 
dyCbo <- dyCbo[ ! grepl("^MT-", rownames(dyCbo)), ]
```

```{r}
# remove all cells with percent.mito > 10%
dyCbo <- subset(dyCbo, subset = percent.mito < 0.1)
dim(dyCbo) # 21,818 genes and 3037 cells
```

```{r}
# save the above object
saveRDS(dyCbo,
        file = "./outputs50K/DAOY-CBO/daoyCbo.seurat.filtCellsGenesNoRiboNoMito.rds")
```

```{r}
# load the above object, which has only been through gene and cell filtering
dyCbo <- readRDS("outputs50K/DAOY-CBO/daoyCbo.seurat.filtCellsGenesNoRiboNoMito.rds")
```

```{r}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference

# include normalisation step of dyCbo before cell cycle scoring as per github issue: https://github.com/satijalab/seurat/issues/1679
dyCbo <- NormalizeData(dyCbo)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
dyCbo <- CellCycleScoring(dyCbo,
                g2m.features = g2m.genes,
                s.features = s.genes)
dyCbo$CC.difference <- dyCbo$S.Score - dyCbo$G2M.Score
```

```{r}
# regressing out nUMI, percent.mito, percent.ribo in addition to CC.difference made cluster assignment
# harder so stick to only regressing CC.difference!!
# this object as input for SCTransform integration, UN-clustered
dyCbo.input.integration <- SCTransform(dyCbo, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```{r}
# save the SCTransformed object, CC difference regressed, PCA reduction, as input for integrated workflow
saveRDS(dyCbo.input.integration, "./outputs50K/DAOY-CBO/DaoyCbo.inputIntegration.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```

```{r DimPlotClusterMapDaoyCbo, fig.height=6, fig.width=10}
# this object is fully SCTransformed and clustered
dyCbo <- SCTransform(dyCbo, vars.to.regress = "CC.difference", method = "glmGamPoi", verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE)
```

```{r}
DimPlot(dyCbo, label = TRUE, repel = TRUE)
```



```{r}
# this object/file is ok to use for DimPlot and other seurat plots, already clustered
saveRDS(dyCbo, file = "./outputs50K/DAOY-CBO/ClusterMapInput.DaoyCbo.NoRiboNoMito.SCT.clustered.CCdiffRegressed.rds")
```

```{r}
dyCbo.markers <- FindAllMarkers(object = dyCbo, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# save the markers of each cluster in DAOY-CBO folder also as RDS file for downstream use
saveRDS(dyCbo.markers, file = "./outputs50K/DAOY-CBO/ClusterMapInput.FindAllMarkers.daoyCbo.rds")
```

```{r}
# save the markers of each cluster in DAOY-CBO folder as csv file
write.csv(dyCbo.markers, file = "./outputs50K/DAOY-CBO/ClusterMapInput.FindAllMarkers.daoyCbo.csv")
```

```{r}
# load the list of DAOY-CBO cluster markers
dyCbo.markers <- read.csv("outputs50K/DAOY-CBO/ClusterMapInput.FindAllMarkers.daoyCbo.csv",
                          stringsAsFactors = FALSE,
                          header = TRUE)
```

```{r}
# look at the top markers of DAOY-CBO clusters
dyCbo.markers %>% 
  group_by(cluster) %>% 
  slice_head(n = 10)
```


****Run ClusterMap - simple run and full run: DAOY-CBO and CBO****
Simple run and full run are performed. ClusterMap seems able to identify the DAOY clusters as clusters 7, 11, 12! Looking at the gene expression for these clusters they appear to match the DAOY clusters I identified manually previously! THE OUTPUT IS AUTOMATICALLY SAVED - SEE OUTPUTS50K/DAOY-CBO/CLUSTERMAP_FILES/dyCboSimple.results.csv
```{r}
# FOR KNIT DO NOT RE-RUN THIS CODE CHUNK
# clusterMap SIMPLE RUN
marker_file_list.dyCbo.simple <- c(cbo = "./outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.csv", 
                      dyCbo = "./outputs50K/DAOY-CBO/ClusterMapInput.FindAllMarkers.daoyCbo.csv")

res.dyCboSimple <- cluster_map(marker_file_list.dyCbo.simple, edge_cutoff = 0.1, output = "dyCboSimple")
res.dyCboSimple
```


```{r}
# all figures are in outputs50K > CBO or DAOY-CBO > figures
# prepare the data
cbo.tsne <- RunTSNE(cbo)
dyCbo.tsne <- RunTSNE(dyCbo)
cbo.dyCbo.combined.tsne <- RunTSNE(cbo.dyCbo.combined)

# save the above TSNE files as RDS files for downstream use
saveRDS(cbo.tsne, file = "./outputs50K/CBO/clusterMap_files/cbo.tsne.CBOsample.ClusterMapInputFullRun.rds")
saveRDS(dyCbo.tsne, file = "./outputs50K/DAOY-CBO/clusterMap_files/DaoyCbo.tsne.DAOY-CBOsample.ClusterMapInputFullRun.rds")
saveRDS(cbo.dyCbo.combined.tsne, file = "./outputs50K/DAOY-CBO/clusterMap_files/DaoyCbo.Cbo.combined.tsne.daoyCboVScboSamples.ClusterMapInputFullRun.rds")
```

```{r}
# clusterMap FULL run - this code produces correct output even though it throws an error message!
marker_file_list <- c(CBO = "outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.csv", 
                      DAOY_CBO = "outputs50K/DAOY-CBO/ClusterMapInput.FindAllMarkers.daoyCbo.csv")

fList <- c(CBO = "outputs50K/CBO/clusterMap_files/cbo.tsne.CBOsample.ClusterMapInputFullRun.rds",
           DAOY_CBO = "outputs50K/DAOY-CBO/clusterMap_files/DaoyCbo.tsne.DAOY-CBOsample.ClusterMapInputFullRun.rds",
           combined = "outputs50K/DAOY-CBO/clusterMap_files/DaoyCbo.Cbo.combined.tsne.daoyCboVScboSamples.ClusterMapInputFullRun.rds")

objList <- lapply(fList, readRDS)

single_obj_list <- 
  c(CBO = objList$CBO, DAOY_CBO = objList$DAOY_CBO)

res.dyCbo.fullRun <- cluster_map(marker_file_list, 
                   edge_cutoff = 0.1, 
                   output = "./outputs50K/DAOY-CBO/figures/combined", 
                   single_obj_list = single_obj_list, 
                   comb_obj = objList$combined)

res.dyCbo.fullRun
```


----------------------------------------------------------------------------------------
****ClusterMap input preparation of the COMBINED CBO and ONS76-CBO object****
Repeat the above with the ONS76-CBO sample
```{r}
# load the ONS76-CBO raw data
data_dir <- "./working_data/ONS76-CBO/"   # the dot is current dir and is essential!
list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r read_expression_matrix}
# note these are .gz files which can be directly read. There are 3654 cells and 36,601 genes.

o76cbo.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                      features = paste0(data_dir, "features.tsv.gz"), 
                      cells = paste0(data_dir, "barcodes.tsv.gz"))
```

```{r}
colnames(cbo.data) = paste0('cbo-', colnames(cbo.data))
colnames(o76cbo.data) = paste0('o76cbo.data-', colnames(o76cbo.data))
```

```{r}
cbo.o76cbo.data = cbind(cbo.data, o76cbo.data)
```

```{r}
cbo.o76cbo.combined <- CreateSeuratObject(counts = cbo.o76cbo.data,
                                         min.cells = 10, min.genes=200,
                                         project = "cbo.o76cbo.comb")

dim(cbo.o76cbo.combined) #23,696 genes and 7294 cells
```

```{r}
# identify mitochondrial genes
mito.genes <- grep(pattern = "^MT", x = rownames(x = cbo.o76cbo.combined), value = TRUE)
percent.mito <- 
  Matrix::colSums(cbo.o76cbo.combined[mito.genes, ])/Matrix::colSums(cbo.o76cbo.combined)

# identify ribosomal genes
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(x = cbo.o76cbo.combined), value = TRUE)
percent.ribo <- 
  Matrix::colSums(cbo.o76cbo.combined[ribo.genes, ])/Matrix::colSums(cbo.o76cbo.combined)
```

```{r}
# add the ribosome and mitochondrial gene percent to the metadata
cbo.o76cbo.combined <- 
  AddMetaData(object = cbo.o76cbo.combined, metadata = percent.mito, col.name = "percent.mito")

cbo.o76cbo.combined <- 
  AddMetaData(object = cbo.o76cbo.combined, metadata = percent.ribo, col.name = "percent.ribo")
```

```{r}
# remove ribosomal and mitochondrial genes
# remove ribosomal genes first
cbo.o76cbo.combined <- cbo.o76cbo.combined[ ! grepl('^RP[SL]', rownames(cbo.o76cbo.combined)), ]

# remove mitochondrial genes 
cbo.o76cbo.combined <- cbo.o76cbo.combined[ ! grepl("^MT-", rownames(cbo.o76cbo.combined)), ]

dim(cbo.o76cbo.combined) # 23,586 genes and 7294 cells
```

```{r}
# remove all cells with percent.mito > 10%
cbo.o76cbo.combined <- subset(cbo.o76cbo.combined, subset = percent.mito < 0.1)

dim(cbo.o76cbo.combined) # 23,586 cells and 5578 genes
```

```{r}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference
# normalise the combined object before the cell cycle scoring and then run SCT (seurat github issue: https://github.com/satijalab/seurat/issues/1679)
cbo.o76cbo.combined <- NormalizeData(cbo.o76cbo.combined)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
cbo.o76cbo.combined <- CellCycleScoring(cbo.o76cbo.combined,
                g2m.features = g2m.genes,
                s.features = s.genes)
cbo.o76cbo.combined$CC.difference <- cbo.o76cbo.combined$S.Score - cbo.o76cbo.combined$G2M.Score
```

```{r}
# start with SCTransform normalisation
cbo.o76cbo.combined <- SCTransform(cbo.o76cbo.combined, 
                   vars.to.regress = "CC.difference", 
                   method = "glmGamPoi", 
                   verbose = FALSE)
```

```{r DimPlotClusterMapCombinedObjectONS76CboAndCbo, fig.height=6, fig.width=10}
# do the clustering
cbo.o76cbo.combined <- cbo.o76cbo.combined %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE)
DimPlot(cbo.o76cbo.combined, label = TRUE, repel = TRUE)
```

```{r}
# save the clustered object
saveRDS(cbo.o76cbo.combined, file = "./outputs50K/ONS76-CBO/clusterMap_files/ClusterMapInput.Ons76Cbo.Cbo.combined.NoRiboNoMito.SCT.clustered.CCdiffRegressed.rds")
```

****ClusterMap input: preparation of the ONS76-CBO object****
```{r}
o76cbo <- CreateSeuratObject(counts = o76cbo.data,
                          min.cells = 10, min.genes=200,
                          project = "ons76cbo")

dim(o76cbo) # 21,213 genes and 3456 cells
```

```{r}
# identify mitochondrial genes
mito.genes <- grep(pattern = "^MT", x = rownames(x = o76cbo), value = TRUE)
percent.mito <- 
  Matrix::colSums(o76cbo[mito.genes, ])/Matrix::colSums(o76cbo)

# identify ribosomal genes
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(x = o76cbo), value = TRUE)
percent.ribo <- 
  Matrix::colSums(o76cbo[ribo.genes, ])/Matrix::colSums(o76cbo)
```

```{r}
# add the ribosome and mitochondrial gene percent to the metadata
o76cbo <- 
  AddMetaData(object = o76cbo, metadata = percent.mito, col.name = "percent.mito")

o76cbo <- 
  AddMetaData(object = o76cbo, metadata = percent.ribo, col.name = "percent.ribo")
```

```{r}
# remove ribosomal and mitochondrial genes
# remove ribosomal genes first
o76cbo <- o76cbo[ ! grepl('^RP[SL]', rownames(o76cbo)), ]

# remove mitochondrial genes 
o76cbo <- o76cbo[ ! grepl("^MT-", rownames(o76cbo)), ]
```

```{r}
# remove all cells with percent.mito > 10%
o76cbo <- subset(o76cbo, subset = percent.mito < 0.1)
dim(o76cbo) # 21,103 genes and 2529 cells
```

```{r}
saveRDS(o76cbo,
        file = "./outputs50K/ONS76-CBO/ons76cbo.seurat.filtCellsGenesNoRiboNoMito.rds")
```

```{r}
o76cbo <- readRDS("outputs50K/ONS76-CBO/ons76cbo.seurat.filtCellsGenesNoRiboNoMito.rds")
```

```{r}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference
# normalise o76cbo before cell cycle scoring as per seurat github issue: https://github.com/satijalab/seurat/issues/1679
o76cbo <- NormalizeData(o76cbo)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
o76cbo <- CellCycleScoring(o76cbo,
                g2m.features = g2m.genes,
                s.features = s.genes)
o76cbo$CC.difference <- o76cbo$S.Score - o76cbo$G2M.Score
```

```{r}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
o76cbo.input.integration <- SCTransform(o76cbo, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```


```{r}
# save the SCTransformed object, CC difference regressed, PCA reduction, as input for integrated workflow
saveRDS(o76cbo.input.integration, "./outputs50K/ONS76-CBO/ons76cbo.inputIntegration.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```


```{r}
o76cbo <- SCTransform(o76cbo, 
                   vars.to.regress = "CC.difference", 
                   method = "glmGamPoi", 
                   verbose = FALSE)
```

```{r DimPlotClusterMapONS76cbo, fig.height=6, fig.width=10}
o76cbo <- o76cbo %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE)
```

```{r}
DimPlot(o76cbo, label = TRUE, repel = TRUE)
```


```{r}
table(o76cbo$seurat_clusters)
```


```{r}
# this file is fine to use for downstream seurat analysis, e.g. violin plots, etc
saveRDS(o76cbo, file = "./outputs50K/ONS76-CBO/ClusterMapInput.ONS76cbo.NoRiboNoMito.SCT.clustered.CCdiffRegressed.rds")
```


```{r}
o76cbo.markers <- FindAllMarkers(object = o76cbo, 
                                 only.pos = TRUE, 
                                 min.pct = 0.25, 
                                 logfc.threshold = 0.25)
```

```{r}
# load the cbo markers to compare with o76cbo markers
cbo.markers <- read.csv("outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.csv",
                        header = TRUE,
                        stringsAsFactors = FALSE)
```

```{r}
# cluster 1 in o76cbo matches quite well to cbo cluster 1, hence the annotation "RP-choroid plexus_like"
# for cluster 1 of o76cbo
cbo.markers %>%
  filter(gene %in% c("SOX2", "IGFBP2", "C1orf61", "ARX", "SOX2-OT", "TTYH1", "HDAC9", "PTCHD1-AS"))
```


```{r}
# look at top markers of ONS76-CBO clusters
o76cbo.markers %>% 
  group_by(cluster) %>% 
  slice_head(n = 30)
```


```{r VlnplotNeuroD1Rbfox3Rtn1ONS76CBOsampleClusterMapInput, fig.width=14}
# this time there are just 2 NEUROD1 clusters (cluster 2 and 3), and cluster 2 is the healthy cluster
VlnPlot(o76cbo, features = c("NEUROD1", "RBFOX3", "RTN1", "CNTN2"))
```

```{r}
# save the markers of each cluster in ONS76-CBO as csv file
write.csv(o76cbo.markers, 
          file = "./outputs50K/ONS76-CBO/ClusterMapInput.FindAllMarkers.ONS76-CBO.csv")
```

```{r}
o76cbo.markers <- read.csv("outputs50K/ONS76-CBO/ClusterMapInput.FindAllMarkers.ONS76-CBO.csv",
                           header = TRUE,
                           stringsAsFactors = FALSE)
```

```{r}
# save the markers of each cluster in ONS76-CBO also as RDS file for downstream use
saveRDS(o76cbo.markers, 
        file = "./outputs50K/ONS76-CBO/ClusterMapInput.FindAllMarkers.ONS76-CBO.rds")
```


****Run ClusterMap - simple run and full run: ONS76-CBO and CBO****
This SIMPLE run result shows that CBO cluster 6 (granule precursors) maps to ONS76-CBO cluster 4, 5 and 11 which are all NEUROD1+ clusters and this matches the ViolinPlot results above (but not NEUROD1+ cluster 9). CBO cluster 7 has no counterpart in ONS76-CBO.
```{r}
# clusterMap SIMPLE RUN
marker_file_list.o76cbo.simple <- c(cbo = "./outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.csv", 
                      o76cbo = "./outputs50K/ONS76-CBO/ClusterMapInput.FindAllMarkers.ONS76-CBO.csv")

res.o76cboSimple <- cluster_map(marker_file_list.o76cbo.simple, 
                                edge_cutoff = 0.1, 
                                output = "o76cboSimple")
res.o76cboSimple
```

ClusterMap FULL RUN:
```{r}
# prepare the data
cbo.tsne <- RunTSNE(cbo)
o76cbo.tsne <- RunTSNE(o76cbo)
cbo.o76cbo.combined.tsne <- RunTSNE(cbo.o76cbo.combined)

# save the above TSNE files as RDS files for downstream use
# saveRDS(cbo.tsne, file = "./outputs50K/CBO/clusterMap_files/cbo.tsne.CBOsample.ClusterMapInputFullRun.rds") (already saved)
saveRDS(o76cbo.tsne, file = "./outputs50K/ONS76-CBO/clusterMap_files/ONS76cbo.tsne.ONS76cboSample.ClusterMapInputFullRun.rds")
saveRDS(cbo.o76cbo.combined.tsne, file = "./outputs50K/ONS76-CBO/clusterMap_files/ONS76cbo.Cbo.combined.tsne.ONS76cboVScboSamples.ClusterMapInputFullRun.rds")
```

```{r}
# results already saved, but figure labels are not ideal!
# clusterMap FULL run - this code produces correct output even though it throws an error message!

marker_file_list <- c(ORG = "outputs50K/CBO/ClusterMapInput.FindAllMarkers.CBO.csv", 
                       ONS76_ORG = "outputs50K/ONS76-CBO/ClusterMapInput.FindAllMarkers.ONS76-CBO.csv")

fList <- c(ORG = "outputs50K/CBO/clusterMap_files/cbo.tsne.CBOsample.ClusterMapInputFullRun.rds",
           ONS76_ORG = "outputs50K/ONS76-CBO/clusterMap_files/ONS76cbo.tsne.ONS76cboSample.ClusterMapInputFullRun.rds",
           combined = "outputs50K/ONS76-CBO/clusterMap_files/ONS76cbo.Cbo.combined.tsne.ONS76cboVScboSamples.ClusterMapInputFullRun.rds")

objList <- lapply(fList, readRDS)

single_obj_list <- 
  c(ORG = objList$ORG, ONS76_ORG = objList$ONS76_ORG)

res.o76cbo.fullRun <- cluster_map(marker_file_list, 
                   edge_cutoff = 0.1, 
                   output = "./outputs50K/ONS76-CBO/figures/combined", 
                   single_obj_list = single_obj_list, 
                   comb_obj = objList$combined)

res.o76cbo.fullRun
```


CODE THAT IS RETAINED AS IT MIGHT BE USEFUL LATER BUT NOT NEEDED CURRENTLY
-----------------------------------------------------------------------------------------

Subset seurat object so that all genes analysed are expressed in at least 10 cells
```
{r}
# this code chunk is not needed, as min.cells=10 in CreateSeuratObject does the same thing
# Only keeping those genes expressed in more than 10 cells
counts <- GetAssayData(object = cbo.dyCbo.combined, slot = "counts")
nonzero <- counts > 0
keep_genes <- Matrix::rowSums(nonzero) >= 10
filtered_counts <- counts[keep_genes, ]
# Reassign to filtered Seurat object
cbo.dyCbo.combined <- CreateSeuratObject(filtered_counts, meta.data = cbo.dyCbo.combined@meta.data)

dim(cbo.dyCbo.combined) # 24,165 genes and 7492 cells
```

```
{r}
# simpler to use min.cells=10 argument in CreateSeuratObject!!!
# Only keeping those genes expressed in more than 10 cells
counts <- GetAssayData(object = cbo, slot = "counts")
nonzero <- counts > 0
keep_genes <- Matrix::rowSums(nonzero) >= 10
filtered_counts <- counts[keep_genes, ]
# Reassign to Seurat object
cbo <- CreateSeuratObject(filtered_counts, meta.data = cbo@meta.data)

dim(cbo) # 21,682 genes and 3838 cells before filtering for mitochondrial percent < 10% and before removing mitochondrial and ribosomal genes
```


***Plot the distribution of proliferative cells in the DimPlot***
Superimpose cycling cells onto the UMAP plot of cell clusters

```
{r}
phase.cbo.sct <- 
  readRDS("outputs50K/CBO/phase.filtRiboMito.cbo.SCT.CCdiffRegressed.clustered.rds")
```

Get the list of genes expressed in cycling cells (Seurat list from Tirosh et al.). This shows that there are much smaller numbers of actively cycling cells than I first assumed.
```
{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
prolif.genes <- list(c(s.genes, g2m.genes))
```


```
{r DimPlotProlifGenes}
phase.cbo.sct <- AddModuleScore(phase.cbo.sct,
                        features = prolif.genes,
                        name = "prolif.genes")
FeaturePlot(phase.cbo.sct, features = 'prolif.genes1', label = TRUE, repel = TRUE) +
   scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu"))) +
  labs()
```

Be careful with the colnames of meta.data. Even though I named "prolif.genes" above, getting the colnames has shown that this becomes "prolif.genes1" !!!
```
{r}
names(x = phase.cbo.sct[[]])
```

Subset the object to extract cluster 5 cells, which are Neurod1+.
```
{r}
neurod1.cbo <- subset(phase.cbo.sct, idents = c("5"))
```

```
{r}
dim(neurod1.cbo)
```

```
# save the neurod1 cluster 5 (which was generated from phase cell cycle scored, CC difference regressed,
# SCTransformed seurat object)
saveRDS(neurod1.cbo, file = "./outputs50K/CBO/neurod1Clust.phaseCCDiffRegress.SCT.cbo.rds")
```

```
save.image("./RDataFiles/CBO_Seurat_EXPLORE1_50K.RData")
```

```
{r}
# Differentially expressed genes using a FOR loop
Idents(phase.cbo.sct.clusters) <- "seurat_clusters"
markers.pos <- data.frame()
for (i in levels(Idents(phase.cbo.sct.clusters))){
  markers <- FindMarkers(phase.cbo.sct.clusters, ident.1 = i, grouping.var = "ident", 
                         verbose = FALSE, only.pos = TRUE, logfc.threshold = 0.25)  # base 2 log fold change
  dat <- markers %>% rownames_to_column(var = "genes")
  dat$cluster_ID <- i
  dat$cluster_ID <- as.numeric(dat$cluster_ID)
  df <- data.frame(dat)
  markers.pos <- rbind(markers.pos, df)
}
head(markers.pos)
```

--------------------------------------------------------------------------------------------
***Repeat clustering of CBO dataset using conventional normalization***
```
# NO NEED TO RUN THESE CODE CHUNKS - SCT transformation worked better

# load the dataset
filt.RiboMito.cbo <- readRDS("outputs50K/CBO/filt.RiboMitoCellsGenes.seurat.cbo.rds")

# load cell cycle genes in advance of regressing out CC difference
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

filt.RiboMito.cbo <- NormalizeData(filt.RiboMito.cbo, normalization.method = "LogNormalize", scale.factor = 10000)
filt.RiboMito.cbo <- FindVariableFeatures(filt.RiboMito.cbo, selection.method = "vst", nfeatures = 2000)

# regress CC difference
filt.RiboMito.cbo <- CellCycleScoring(phase.cbo.sct,
                g2m.features = g2m.genes,
                s.features = s.genes)
filt.RiboMito.cbo$CC.difference <- filt.RiboMito.cbo$S.Score - filt.RiboMito.cbo$G2M.Score

filt.RiboMito.cbo <- ScaleData(filt.RiboMito.cbo, vars.to.regress = "CC.difference")

# run the rest of the seurat workflow
filt.RiboMito.cbo <- RunPCA(filt.RiboMito.cbo, features = VariableFeatures(object = filt.RiboMito.cbo))
filt.RiboMito.cbo <- FindNeighbors(filt.RiboMito.cbo, dims = 1:10)
filt.RiboMito.cbo <- FindClusters(filt.RiboMito.cbo, resolution = 0.8) # this res gives only 1 NeuroD1 cluster

filt.RiboMito.cbo <- RunUMAP(filt.RiboMito.cbo, dims = 1:10)
DimPlot(filt.RiboMito.cbo, reduction = "umap")
```

```
VlnPlot(filt.RiboMito.cbo, features = c("NEUROD1", "RBFOX3", "RTN1"))
```

This code retained just to show arguments to read a csv file
```
# load 'cbo.cluster.markers.NoRiboNoMito.csv'

cbo.cluster.markers.noRiboMito <- read.csv("outputs50K/CBO/cbo.cluster.markers.NoRiboNoMito.csv", 
                                       header = TRUE, 
                                       stringsAsFactors = FALSE)
```

This code is retained just to show the usage of write.csv
```
# save cbo.cluster.markers.ribo20 as a csv file. This object has percent.ribo < 20

write.csv(cbo.cluster.markers.ribo20, "./outputs50K/CBO/cbo.cluster.markers.ribo20.csv", quote = F)
```

The DotPlots are not that useful
```
{r DotplotSeppTFmarkers}
# using the Sepp et al paper (Extended Fig 8g), test if markers from that paper are expressed 
# in the seurat object filtered to remove all ribosomal and mitochondrial genes, and where cells which
# have mitoch genes > 10% are excluded.
# MUST use the already clustered seurat object.
# I used just the transcription factor (TF) markers (sepp.TF.markers)

sepp.TF.markers <- c("PAX3", "RFX4", "TCFL71", "GLIS3", "SOX6", "TCF7L2", "HOPX", "TSC22D4", "ID4",
             "ESRRB", "FOXP2", "FOXP4", "MEF2C",
             "PAX2", "TFAP2A", "TFAP2B", "BHLHE22", "PRDM8",
             "PAX6", "BARHL1", "E2F3", "ST18", "SATB2", "ZBTB18", "NEUROD1", "ETV1", "KLF9", "NR3C1")


DotPlot(object = phase.cbo.sct.clusters, features = sepp.TF.markers) +
  coord_flip()
```

```
{r DotplotSeppMarkersAndCiliaManyCBO, fig.height=10}
# from the Sepp paper repeat Dotplot with TF and some key non-TF markers
# added markers of ciliated cells from cluster10 above.
# used the argument col.min = 0 so that average expression scale is from 0 to 2.5 in key

sepp.all.markers <- c("TFAP2A", "PAX3", "RFX4", "TCFL71", "TCF7L2", "GLIS3", "SOX6", "HOPX", "TSC22D4",  
                      "ID4", "NCKAP5", "SLIT2", "CLYBL", "SLC1A3", "SLC6A11", "AQP4",  #astroglial/progenitor
                      
                      "ESRRB", "FOXP1", "FOXP2", "FOXP4", "MEF2C", "RORB", "CDH9", "ETV1", "EBF1", "EBF2", 
                      "SKOR2", "DAB1", "IPTR",   # Purkinje
                      
                      "NOTCH1",    # VZ neuroblast
             
                      "PAX2", "NFIA", "NFIB", "TFAP2B", "BHLHE22", "PRDM8",  # interneuron
                      
                      "ZFHX3", "ZFHX4",   # GABAergic deep nuclei neurons
             
                      "TOP2A", "ATOH1", "LMX1A", "OTX2","ETV6", "PAX6", "BARHL1", "E2F3", "ST18", "SATB2", 
                      "ZBTB18", "NEUROD1", "ZIC1", "ZIC4", "GABAR6", "CNTN1", "CNTN2", "KLF9", "NR3C1", "LMO3", 
                      "SYNPR", "KCNIP4", "HCRTR2", 
                      "EOMES", "TRPC3", "GRM1")
                      # UBC/GCP/granule/glut deep cerebellar nuclei                                                                   
seppAll.cilia.stem.markers <- c(sepp.all.markers, "CCNO", "CFAP299", "ROPN1L", "C11orf88", "ANKRD66", "C20orf85", 
                                "STOML3", "TEKT4", "SNTN", "MEIS1", "POU3F2", "HOXB2", "MEIS2", "GRIA4")

DotPlot(object = phase.cbo.sct.clusters, features = seppAll.cilia.stem.markers, col.min = 0) +
  coord_flip()
```







