---
title: "DAOY-2Drep_Seurat_EXPLORE1_50K"
author: "JJ"
date: "2023-01-26"
output: html_document
---

```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE,
                      fig.path = "./outputs50K/DAOY_2D/figures/")
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(glmGamPoi)
library(presto)
library(msigdbr)
library(fgsea)
library(RColorBrewer)
library(org.Hs.eg.db)
library(DOSE)
library(clustree)
library(tidyverse)
library(magrittr)
library(patchwork)
})
```

```{r eval=FALSE, echo=FALSE}
# load the data from scratch
# load the DAOY-2D raw data (from "working_data" folder)
data_dir <- "./working_data/DAOY-2D/"   # the dot is current dir and is essential!
list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r eval=FALSE, echo=FALSE}
# note these are .gz files which can be directly read. n = 2922 cells (dyMono.data)
dyMono.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                         features = paste0(data_dir, "features.tsv.gz"), 
                         cells = paste0(data_dir, "barcodes.tsv.gz"))
# create the seurat object, keep genes expressed in 10 or more cells
dyMono <- CreateSeuratObject(counts = dyMono.data,
                          min.cells = 10, min.genes=2000,  # set high as this sample was very deeply sequenced
                          project = "dyMono")

dim(dyMono) # 24,957 genes and 2922 cells
```

```{r eval=FALSE, echo=FALSE}
# identify mitochondrial genes
mito.genes <- grep(pattern = "^MT", x = rownames(x = dyMono), value = TRUE)
percent.mito <- 
  Matrix::colSums(dyMono[mito.genes, ])/Matrix::colSums(dyMono)

# add the mitochondrial gene percent to the metadata
dyMono <- 
  AddMetaData(object = dyMono, metadata = percent.mito, col.name = "percent.mito")

# remove mitochondrial genes 
dyMono <- dyMono[ ! grepl("^MT-", rownames(dyMono)), ]

dim(dyMono) # 24,944 genes and 2922 cells
```

```{r eval=FALSE, echo=FALSE}
# remove all cells with percent.mito > 10%
dyMono <- subset(dyMono, subset = percent.mito < 0.1)
dim(dyMono) # reduced to 24,944 genes and 2633 cells
```

```{r eval=FALSE, echo=FALSE}
# save the NON-normalised, NON-clustered seurat object
saveRDS(dyMono, file = "./outputs50K/DAOY_2D/filt.MitoCellsGenes.RiboRetained.seurat.DaoyMonolayer.rds")
```

```{r eval=FALSE, echo=FALSE}
dyMono <- readRDS("outputs50K/DAOY_2D/filt.MitoCellsGenes.RiboRetained.seurat.DaoyMonolayer.rds")
```

```{r eval=FALSE, echo=FALSE}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference

# normalise object before running cell cycle scores as per Seurat github issue: https://github.com/satijalab/seurat/issues/1679
dyMono <- NormalizeData(dyMono)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
dyMono <- CellCycleScoring(dyMono,
                g2m.features = g2m.genes,
                s.features = s.genes)
dyMono$CC.difference <- dyMono$S.Score - dyMono$G2M.Score
```

```{r eval=FALSE, echo=FALSE}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
dyMono.input.integration <- SCTransform(dyMono, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
# saved the SCTransformed object, CC difference regressed, PCA reduction, as input for integrated workflow
saveRDS(dyMono.input.integration, "./outputs50K/DAOY_2D/DAOYmono.inputIntegration.RiboRetained.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```

```{r eval=FALSE, echo=FALSE}
dyMono <- SCTransform(dyMono, vars.to.regress = "CC.difference", method = "glmGamPoi", verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.5)
```

```{r eval=FALSE, echo=FALSE}
# save the clustered, SCTransformed object above
saveRDS(dyMono, file = "./outputs50K/DAOY_2D/DAOYmono.MitoCellsGenesFilt.RiboRetained.SCT.clustered.CCdiffRegressed.rds")
```

```{r}
# load the above object.
dyMono <- readRDS("outputs50K/DAOY_2D/DAOYmono.MitoCellsGenesFilt.RiboRetained.SCT.clustered.CCdiffRegressed.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotDAOYmonoRepUmapSCTransform, fig.height=6, fig.width=10}
DimPlot(dyMono, label = TRUE, pt.size = 1, label.size = 4)
```

```{r eval=FALSE, echo=FALSE}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
prolif.genes <- list(c(s.genes, g2m.genes))
```


```{r eval=FALSE, echo=FALSE, AddModuleScoreProlifernDAOYmonolayer, fig.height=6, fig.width=10}
phase.dyMono <- AddModuleScore(dyMono,
                        features = prolif.genes,
                        name = "prolif.genes")
FeaturePlot(phase.dyMono, features = 'prolif.genes1', label = TRUE, repel = TRUE, pt.size = 1,
            label.size = 4) +
   scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu"))) +
  labs(title = "")
```

```{r eval=FALSE, echo=FALSE, VlnPlotGranuleGenesDAOYmono, fig.height=17, fig.width=22}
p1a <- VlnPlot(dyMono,
               features = "MEIS1") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1b <- VlnPlot(dyMono,
               features = "MEIS2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1c <- VlnPlot(dyMono,
               features = "SATB2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1d <- VlnPlot(dyMono,
               features = "ZIC1") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1e <- VlnPlot(dyMono,
               features = "ZIC2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1f <- VlnPlot(dyMono,
              features = "JKAMP") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1a + p1b + p1c + p1d + p1e + p1f + plot_layout(ncol = 2)
```

```{r eval=FALSE, echo=FALSE}
VlnPlot(dySpher, features = c("CNTN1", "CNTN2"))
```



```{r eval=FALSE, echo=FALSE}
# get the markers of each cluster - slow to run
dyMono.POSmarkers <- 
  FindAllMarkers(object = dyMono, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

Cluster 5 is enriched for ribosome genes
```{r eval=FALSE, echo=FALSE}
dyMono.POSmarkers %>% 
  dplyr::group_by(cluster) %>% 
  slice_head(n = 30)
```

```{r eval=FALSE, echo=FALSE}
# None of these five genes is expressed
dyMono.POSmarkers %>%
  filter(gene %in% c("NEUROD1", "CNTN1", "CNTN2", "MEIS1", "DCX"))
```

```{r eval=FALSE, echo=FALSE}
# save the positive markers of DAOY-2D (rep)
saveRDS(dyMono.POSmarkers, 
        file = "./outputs50K/DAOY_2D/POS.markers.RiboRetained.DAOYmonolayerRep.SCT.clustered.rds")
```

```{r eval=FALSE, echo=FALSE}
dyMono.POSmarkers <- readRDS("outputs50K/DAOY_2D/POS.markers.RiboRetained.DAOYmonolayerRep.SCT.clustered.rds")
```

```{r eval=FALSE, echo=FALSE}
# first load dyMono seurat object above, then only run this chunk
dyMono.POSmarkers %>% 
  group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
```

```{r eval=FALSE, echo=FALSE}
# this step is important to avoid a warning re omission of specific genes of interest in the top 10 list of markers from the scale data slot
# see https://github.com/satijalab/seurat/issues/1369
dyMono <- ScaleData(object = dyMono, features = rownames(dyMono)) 
```

```{r eval=FALSE, echo=FALSE, HeatmapDaoyMonoTop10markers5clustersRes0.5, fig.height=8, fig.width=10}
DoHeatmap(dyMono, features = top10$gene) + guides(color = "none") +
  theme(text = element_text(size = 25), legend.title = element_text(size=20), 
        legend.text=element_text(size=20))
```

***FGSEA cell clusters DAOY-monolayer (rep)***
```{r eval=FALSE, echo=FALSE}
# prepare the data for gProfiler (web app) input
dyMono.genes.presto <- wilcoxauc(dyMono, 'seurat_clusters')
nrow(dyMono.genes.presto)
dplyr::count(dyMono.genes.presto, group)
```

****FGSEA cell clusters****
```{r eval=FALSE, echo=FALSE}
# use dplyr to get the appropriate gene set
msig.hs <- msigdbr(species = "Homo sapiens")

msig.GOBP.hs <- msig.hs %>% 
dplyr::filter(gs_subcat == "GO:BP")

head(msig.GOBP.hs)
```

```{r eval=FALSE, echo=FALSE}
fgsea_sets_GOBP <- msig.GOBP.hs %>% split(x = .$gene_symbol, f = .$gs_name)
```


*****Cluster 0 FGSEA*****
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.genes.presto.clust0 <- dyMono.genes.presto %>% 
  dplyr::filter(group == "0" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.genes.presto.clust0) # 9 genes only!
```

```{r eval=FALSE, echo=FALSE}
ranks.clus0 <- deframe(dyMono.genes.presto.clust0)
head(ranks.clus0)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus0.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus0, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - NO enriched pathways!
fgseaResTidy.clus0.GOBP <- fgseaRes.clus0.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus0.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) # there are no enriched pathways in this cluster, nothing to save!
```

*****Cluster 1 FGSEA*****
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.genes.presto.clust1 <- dyMono.genes.presto %>% 
  dplyr::filter(group == "1" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.genes.presto.clust1) # 658 genes
```

```{r eval=FALSE, echo=FALSE}
ranks.clus1 <- deframe(dyMono.genes.presto.clust1)
head(ranks.clus1)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus1.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus1, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - NO enriched pathways!
fgseaResTidy.clus1.GOBP <- fgseaRes.clus1.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus1.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) # there are no enriched pathways in this cluster, nothing to save!
```

```{r eval=FALSE, echo=FALSE}
# use fwrite() as a column of the tibble is a list!
fwrite(fgseaResTidy.clus1.GOBP, 
       file = "./outputs50K/DAOY_2D/cluster1.DAOYmono.multifgsea.presto.GOBiolProcessGenes.csv")
```


*****Cluster 2 FGSEA*****
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.genes.presto.clust2 <- dyMono.genes.presto %>% 
  dplyr::filter(group == "2" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.genes.presto.clust2) # 555 genes
```

```{r eval=FALSE, echo=FALSE}
ranks.clus2 <- deframe(dyMono.genes.presto.clust2)
head(ranks.clus2)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus2.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus2, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - NO enriched pathways!
fgseaResTidy.clus2.GOBP <- fgseaRes.clus2.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus2.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) # there are no enriched pathways in this cluster, nothing to save!
```

*****Cluster 3 FGSEA*****
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.genes.presto.clust3 <- dyMono.genes.presto %>% 
  dplyr::filter(group == "3" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.genes.presto.clust3) # 570 genes
```

```{r eval=FALSE, echo=FALSE}
ranks.clus3 <- deframe(dyMono.genes.presto.clust3)
head(ranks.clus3)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus3.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus3, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - NO enriched pathways!
fgseaResTidy.clus3.GOBP <- fgseaRes.clus3.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus3.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) # there are no enriched pathways in this cluster, nothing to save!
```

```{r eval=FALSE, echo=FALSE}
# use fwrite() as a column of the tibble is a list!
fwrite(fgseaResTidy.clus3.GOBP, 
       file = "./outputs50K/DAOY_2D/cluster3.DAOYmono.multifgsea.presto.GOBiolProcessGenes.csv")
```


*****Cluster 4 FGSEA*****
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.genes.presto.clust4 <- dyMono.genes.presto %>% 
  dplyr::filter(group == "4" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.genes.presto.clust4) # 803 genes
```

```{r eval=FALSE, echo=FALSE}
ranks.clus4 <- deframe(dyMono.genes.presto.clust4)
head(ranks.clus4)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus4.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus4, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - NO enriched pathways!
fgseaResTidy.clus4.GOBP <- fgseaRes.clus4.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus4.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) # there are no enriched pathways in this cluster, nothing to save!
```


*****Cluster 5 FGSEA*****
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.genes.presto.clust5 <- dyMono.genes.presto %>% 
  dplyr::filter(group == "5" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.genes.presto.clust5) # 750 genes
```

```{r eval=FALSE, echo=FALSE}
ranks.clus5 <- deframe(dyMono.genes.presto.clust5)
head(ranks.clus5)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus5.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus5, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - NO enriched pathways!
fgseaResTidy.clus5.GOBP <- fgseaRes.clus5.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus5.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) # there are no enriched pathways in this cluster, nothing to save!
```

```{r eval=FALSE, echo=FALSE}
# use fwrite() as a column of the tibble is a list!
fwrite(fgseaResTidy.clus5.GOBP, 
       file = "./outputs50K/DAOY_2D/cluster5.DAOYmono.multifgsea.presto.GOBiolProcessGenes.csv")
```









As can be seen from the above heatmap the res=0.5 clustering seems to generate too many clusters. For example, cluster 2 has many genes in common with cluster 1. Also cluster 3 and cluster 5 share many markers in common. Try resolution=0.1 and make a new heatmap - res = 0.1 chosen as in the integration of DAOY-2D first sample and DAOY-2D repeat sample, there were only 2 major clusters

```{r eval=FALSE, echo=FALSE}
dyMono.res0.1 <- SCTransform(dyMono, vars.to.regress = "CC.difference", method = "glmGamPoi", verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.1)
```

```{r eval=FALSE, echo=FALSE, DimplotDAOYmonoRepUmapSCTransformRes0.1, fig.height=6, fig.width=10}
DimPlot(dyMono.res0.1, label = TRUE)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(dyMono.res0.1, 
        file = "./outputs50K/DAOY_2D/DAOYmono.MitoCellsGenesFilt.RiboRetained.SCT.Res0.1.clustered.CCdiffRegressed.rds")
```


```{r eval=FALSE, echo=FALSE}
# get the markers of each cluster - slow to run
dyMono.res0.1.POSmarkers <- 
  FindAllMarkers(object = dyMono.res0.1, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r eval=FALSE, echo=FALSE}
dyMono.res0.1.POSmarkers %>%
  dplyr::group_by(cluster) %>% 
  slice_head(n = 20)
```

```{r eval=FALSE, echo=FALSE}
# this clustering is better than res=0.5 clustering which over-clusters the sample
saveRDS(dyMono.res0.1.POSmarkers,
        file = "./outputs50K/DAOY_2D/POS.markers.RiboRetained.DAOYmonolayerRep.SCT.Res0.1.clustered.rds")
```

```{r eval=FALSE, echo=FALSE}
dyMono.res0.1.POSmarkers <- readRDS("outputs50K/DAOY_2D/POS.markers.RiboRetained.DAOYmonolayerRep.SCT.Res0.1.clustered.rds")
```

```{r eval=FALSE, echo=FALSE}
dyMono.res0.1.POSmarkers %>% 
  group_by(cluster) %>%
    top_n(n = 20, wt = avg_log2FC) -> top20
```

```{r eval=FALSE, echo=FALSE}
# this step is important to avoid a warning re omission of specific genes of interest in the top 10 list of markers from the scale data slot
# see https://github.com/satijalab/seurat/issues/1369
dyMono.res0.1.scaled <- ScaleData(object = dyMono.res0.1, features = rownames(dyMono.res0.1)) 
```

Note that cluster 2 genes visually have the least expression difference compared to cluster 0 and cluster 1
```{r eval=FALSE, echo=FALSE, HeatmapDaoyMonoTop20markers5clustersRes0.1, fig.height=8, fig.width=10}
DoHeatmap(dyMono.res0.1.scaled, features = top20$gene) + guides(color = "none") +
  theme(text = element_text(size = 25), legend.title = element_text(size=20), 
        legend.text=element_text(size=20))
```

Run PRESTO
```{r eval=FALSE, echo=FALSE}
# prepare the data for gProfiler (web app) input
dyMono.res0.1.genes.presto <- wilcoxauc(dyMono.res0.1, 'seurat_clusters')
nrow(dyMono.res0.1.genes.presto)
dplyr::count(dyMono.res0.1.genes.presto, group)
```

```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.res0.1.genes.presto.clust0 <- dyMono.res0.1.genes.presto %>% 
  dplyr::filter(group == "0" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.res0.1.genes.presto.clust0) # 1131 genes
```

```{r eval=FALSE, echo=FALSE}
# gPROFILER input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(dyMono.res0.1.genes.presto.clust0, 
         "./outputs50K/DAOY_2D/daoyMono.res0.1.cluster0.genes.RiboRetained.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.res0.1.genes.presto.clust1 <- dyMono.res0.1.genes.presto %>% 
  dplyr::filter(group == "1" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.res0.1.genes.presto.clust1) # 550 genes
```

```{r eval=FALSE, echo=FALSE}
# gPROFILER input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(dyMono.res0.1.genes.presto.clust1, 
          "./outputs50K/DAOY_2D/daoyMono.res0.1.cluster1.genes.RiboRetained.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
# filter and order the gene list
dyMono.res0.1.genes.presto.clust2 <- dyMono.res0.1.genes.presto %>% 
  dplyr::filter(group == "2" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.res0.1.genes.presto.clust2) # 9 genes only, WHICH ARE NOT WORTH SAVING!
```


***Clustree***
```{r eval=FALSE, echo=FALSE}
# attempt to use Clustree package to find the right number of clusters
# -----------------------------------------------------------------------------------------------

# Select a range of resolutions
resolution.range <- seq(from = 0, to = 1.8, by = 0.2)

# Find clusters using a range of resolutions, the clustering information is in the metadata slot
dyMono.clusters <- FindClusters(object = dyMono.res0.2, 
                                      resolution = resolution.range,
                                      verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE, ClustreeCBO, fig.height=8}
# the clustering tree finds too many clusters.
# with res = 0.8 there are 4 clusters; with res = 1 there are 5 clusters.
# use marker differences (FindAllMarkers) and pathway enrichment to tell how many clusters have distinct enrichments. 

clustree(dyMono.clusters)
```

```{r eval=FALSE, echo=FALSE}
# get the number of cells per cluster
table(Idents(dyMono.clusters))
```







