---
title: "DAOY-2D_rep_Seurat_QC1_50K"
author: "JJ"
date: "02/10/2022"
output: html_document
---
Sequencing was done on the repeat DAOY-2D sample, which was a very high number of reads on 2922 cells
```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE)
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(glmGamPoi)
library(Matrix)
library(patchwork)
library(tidyverse)
})
```

```{r path_to_data}
data_dir <- "./working_data/DAOY-2D/"   # the dot is essential!

list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r read_expression_matrix}
# note these are .gz files which can be directly read. The number of barcodes detected
# is 2922 and there are 36,601 genes

daoy.2d.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                         features = paste0(data_dir, "features.tsv.gz"), 
                         cells = paste0(data_dir, "barcodes.tsv.gz"))

dim(daoy.2d.data)
```

```{r distrib_detect_genes_stats}
# Check how many genes have at least one transcript in each cell. 

at_least_one <- apply(daoy.2d.data, 2, function(x) sum(x>0)) # counts the sum of rows that have a                                                                         count value (x > 0) for each column
summary(at_least_one) # median number of genes is higher than for 20K dataset

# summary of total expression per cell. It's high (median total expressn=61,834 genes/cell), which is just reflecting that very deep sequencing was performed on the repeat sample.
summary(colSums(daoy.2d.data))
```

```{r Plots_GenesPerCell, fig.height=10, fig.width=8}

par(mfrow = c(2, 1))

hist(at_least_one, breaks = 100,
     main = "Distribution of detected genes",
     xlab = "Genes with at least one transcript")

hist(colSums(daoy.2d.data),                             # total counts (expression) per cell
     breaks = 100, main = "Expression sum per cell",
     xlab = "Sum expression")

```

```{r genesPerCell}
# manually check the number of genes (that have transcripts) detected in three or more cells
# ~28.4K genes are expressed in at least 3 cells
# ~8.1K genes are not expressed in at least 3 cells
tmp <- apply(daoy.2d.data, 1, function(x) sum(x > 0))
table(tmp >= 3)
length(tmp) # total number of genes expressed by all cells in dataset
```

```{r cells_200genes}
# NOT all 2922 cells have at least 200 (common Seurat threshold) expressed genes
# (min = 363 genes, see below)

keep <- tmp>=3                  # for each gene this states TRUE or FALSE

tmp2 <- daoy.2d.data[keep, ]   # subset ons76.2d.data to keep only TRUE values, i.e. genes which 
                                # are expressed in 3 or more cells

dim(tmp2)                       # dim is 19,194 by 2922, as, for example for some genes the 2922th cell may
                                # be one of the cells expressing that gene

genes_in_3orMoreCells <- apply(tmp2, 2, function(x) sum(x > 0))   # for each cell, this counts the number 
                                                                  # of genes that cell expresses for all
                                                                  # genes expressed in 3 or more cells

summary(genes_in_3orMoreCells)
```

```{r Plot_sorted_genesPerCell}
# shape of graph is different from expected. Appears to be an elbow at sorted genes per cell=2000
genes_per_cell <- Matrix::colSums(daoy.2d.data > 0) # count gene only if it has non-zero reads                                                             mapped.

plot(sort(genes_per_cell), 
     xlab='cell',
     ylab = "sorted genes per cell",
     log = "y",                             # only y axis is logarithmic
     main='Genes per cell (ordered)') 
```

```{r create_seurat_object}
# no. of genes has dropped to 28,396 when min.cells = 3, and set min.featres = 2000

daoy.2d.seurat <- CreateSeuratObject(counts = daoy.2d.data, 
                                    min.cells = 3, 
                                    min.features = 2000, # set this number high as sequencing depth was v. high
                                    project = "DAOY_2Drep")
dim(daoy.2d.seurat)
```

```{r add_percent_mitochondria_meta.data}
# add mitochondrial percentage to metadata
# the 'orig.ident' field is labelled based on the CreateSeuratObject(project=) function. So this field can
# be changed by going back into the CreateSeuratObject () function and making the change to that command.
daoy.2d.seurat[["percent.mt"]] <- PercentageFeatureSet(daoy.2d.seurat, pattern = "^MT-")
head(daoy.2d.seurat@meta.data)
```

```{r metadata_dataframe_2922cells}
# create metadata dataframe on its own
metadata <- daoy.2d.seurat@meta.data

# Add cell IDs to metadata
metadata$cells <- rownames(metadata)

# Rename columns to be more intuitive
metadata <- metadata %>%
        dplyr::rename(sample = orig.ident,
                      nUMI = nCount_RNA,
                      nGene = nFeature_RNA)
dim(metadata)
```

```{r violinPlot_genes_transcripts_percent.mt_2120cells, fig.height=10, fig.width=5}
# there are some cells with excessive mitochondrial transcripts
# the data looks bimodal, smaller lower 'bulge'

p1 <- metadata %>% 
        ggplot(aes(x = sample, y = nUMI)) +
        geom_violin(colour = "blue", fill = "blue", alpha = 0.1) +
        theme_classic() +
        geom_jitter(alpha = 0.2)

p2 <- metadata %>% 
        ggplot(aes(x = sample, y = nGene)) +
        geom_violin(colour = "brown1", fill = "brown1", alpha = 0.1) +
        theme_classic() +
        geom_jitter(alpha = 0.2)

p3 <- metadata %>% 
        ggplot(aes(x = sample, y = percent.mt)) +
        geom_violin(colour = "cadetblue", fill = "cadetblue", alpha = 0.1) +
        theme_classic() +
        geom_jitter(alpha = 0.2)

p1 / p2 / p3
```

```{r Plot_UMIcountsPerCell_2922cells}
# visualise number of UMIs (transcripts) per cell using all 2922 cells. 
# In geom_density use (y = ..scaled..) value that stat_density provides

metadata %>% 
        ggplot(aes(x = nUMI)) +
        geom_density(aes(y = ..scaled..)) +
        scale_x_log10() +
        theme_classic() +
        ylab("cell density") +
        geom_vline(xintercept = 500, color = "blue")
```

```{r mitochondrial_thresholds}
# number of cells with percent.mt > 10%: 282 cells
# number of cells with percent.mt < 10%: 2640 cells

nrow(metadata[metadata$percent.mt > 10, ])

nrow(metadata[metadata$percent.mt < 10, ])
```

```{r filtered_mitoch_seuratObject}
# remove cells with percent mitochondrial gene expression > 10, i.e. remove 185 cells
# CELL-LEVEL filtering leaves 1935 cells

filtered_daoy.2d.seurat <- subset(daoy.2d.seurat, subset = nFeature_RNA > 200 & percent.mt < 10)
dim(filtered_daoy.2d.seurat)
```

```{r violinPlot_genes_transcripts_percent.mt1935cells, fig.height=10, fig.width=5}
# create a metadata dataframe
# new plots show remaining data looks unimodal
# mitochondrial gene filtration successful

metadata.2 <- filtered_daoy.2d.seurat@meta.data

# Add cell IDs to metadata
metadata.2$cells <- rownames(metadata.2)

# Rename columns to be more intuitive
metadata.2 <- metadata.2 %>%
        dplyr::rename(sample = orig.ident,
                      nUMI = nCount_RNA,
                      nGene = nFeature_RNA)

p4 <- metadata.2 %>% 
        ggplot(aes(x = sample, y = nUMI)) +
        geom_violin(colour = "blue", fill = "blue", alpha = 0.1) +
        theme_classic() +
        geom_jitter(alpha = 0.1)

p5 <- metadata.2 %>% 
        ggplot(aes(x = sample, y = nGene)) +
        geom_violin(colour = "brown1", fill = "brown1", alpha = 0.1) +
        theme_classic() +
        geom_jitter(alpha = 0.1)

p6 <- metadata.2 %>% 
        ggplot(aes(x = sample, y = percent.mt)) +
        ylim(0, 100) +
        geom_violin(colour = "cadetblue", fill = "cadetblue", alpha = 0.1) +
        theme_classic() +
        geom_jitter(alpha = 0.1)

p4 / p5 / p6
```

```{r extract_countsSlot1935cells}
# subset '@counts' slot (matrix) of filtered seurat object
counts.1935 <- GetAssayData(object = filtered_daoy.2d.seurat, slot = "counts")
```

```{r summarise_stats1935cells}
# median genes per cell and expression sum, as above. Min. genes = 561. Median = 9526, Max = 15K genes!
# Median genes is higher than before cells with high mitoch reads filtered out, which makes sense as the latter cells typically have lower number of genes expressed

at_least_one.1935 <- apply(counts.1632, 2, function(x) sum(x>0))

summary(at_least_one.1935)

summary(colSums(counts.1935))
```

```{r Plots_GenesPerCell1935cells, fig.height=10, fig.width=8}
# 1 major peak now, but still quite a large group of cells to the left with lower gene expression,
# and there is a tail on the right for cells expressing v. large number of genes.
# In the bottom graph far fewer cells with low expression sum

par(mfrow = c(2, 1))

hist(at_least_one.1935, breaks = 100,
     main = "Distribution of detected genes",
     xlab = "Genes with at least one transcript")

hist(colSums(counts.1935),                             # total counts (expression) per cell
     breaks = 100, main = "Expression sum per cell",
     xlab = "Sum expression")
```

```{r Plots_UMI_gene_expression_1935cells, fig.height=10, fig.width=5}

# unimodal plots with cells expressing 3e+04 (30,000) to 1e+05 (100,000) UMIs and around 10,000 genes

p7 <- metadata.2 %>% 
       ggplot(aes(x = nUMI)) +
        geom_density(aes(y = ..scaled..), colour = "blue", fill = "blue", alpha = 0.2) +
        theme_classic() +
        scale_x_log10() +
        ylab("Cell density")

p8 <- metadata.2 %>% 
        ggplot(aes(x = nGene)) +
        geom_density(aes(y = ..scaled..), colour = "brown1", fill = "brown1", alpha = 0.2) +
        theme_classic() +
        scale_x_log10() +
        ylab("Cell density")

p7 / p8
```

```{r Plot_UMI_v_genes_1935cells}
# plot of UMI count vs no. of genes

p9 <- FeatureScatter(filtered_daoy.2d.seurat, 
                      feature1 = "nCount_RNA",        # nCount_RNA is same as UMI count
                      feature2 = "nFeature_RNA")

p9
```

```{r PlotSortedGenesPerCell1935cells}
# shape of graph is now as expected
genes_per_cell <- Matrix::colSums(filtered_daoy.2d.seurat) # count gene only if it has non-zero reads                                                             mapped.

plot(sort(genes_per_cell), 
     xlab='cell',
     ylab = "sorted genes per cell",
     log = "y",                             # only y axis is logarithmic
     main='Genes per cell (ordered)') 
```


```{r filtered_seuratObject_GeneLevel}
# GENE LEVEL FILTERING (applied on cell-filtered Seurat object)
#-----------------------------------------------------------------------------------
#-----------------------------------------------------------------------------------
#  we will have many genes with zero counts, which can dramatically reduce the average expression for a cell and so we will remove them from our data. Remove genes that have zero expression in all cells.

# # Extract counts
counts <- GetAssayData(object = filtered_daoy.2d.seurat, slot = "counts")

# Output a logical vector for every gene on whether there are more than zero counts per cell
nonzero <- counts > 0

# If a gene is only expressed in a handful of cells, it is not particularly meaningful as it still brings down the averages for all other cells it is not expressed in. For our data we choose to keep only genes which are expressed in 10 or more cells.
# Sums all TRUE values and returns TRUE if more than 10 TRUE values per gene
keep_genes <- Matrix::rowSums(nonzero) >= 10    # sum the TRUE values in a row with 'gene' as rowname

# Only keeping those genes expressed in more than 10 cells
filtered_counts <- counts[keep_genes, ]

# Reassign to filtered Seurat object
filtCellsGenes.daoy.2d.seurat <- CreateSeuratObject(filtered_counts, meta.data = filtered_daoy.2d.seurat@meta.data)

dim(filtCellsGenes.daoy.2d.seurat)       # number of genes further reduced to ~ 24.7K genes from ~ 28.4K genes                                                        for only cell-level filtering
```

```{r}
# saved the 'filtCellsGenes.ons76.2d.seurat' object to outputs50K/ONS76_2D folder
saveRDS(filtCellsGenes.daoy.2d.seurat, file = "./outputs50K/DAOY_2D/filtCellsGenes.daoy.2d.seurat.rds")
```

--------------------------------------------------------------------------------------------
***Filter the above Seurat object further to remove mitochondrial and ribosomal genes***

```{r}
filtCellsGenes.daoy.2d.seurat <- readRDS("outputs50K/DAOY_2D/filtCellsGenes.daoy.2d.seurat.rds")
```

```{r removeRiboGenes}
# FILTER OUT ALL RIBOSOMAL GENES, THEN ALL MITOCHONDRIAL GENES (ON THE CELLS WHICH HAVE ALREADY BEEN REMOVED
# THAT HAVE PERCENT.MITO < 10% BEFORE DOING THE CLUSTERING), DON'T REMOVE ANY MORE ACTUAL CELLS
#-------------------------------------------------------------------------------------

# remove ribosomal genes first
filt.CellsGenesRibo.daoy2d <- filtCellsGenes.daoy.2d.seurat[ ! grepl('^RP[SL]', 
                                                                     rownames(filtCellsGenes.daoy.2d.seurat)), ]

# removal of ribosomal genes works - from 24,726 genes down to 24,628 genes
dim(filtCellsGenes.daoy.2d.seurat)
dim(filt.CellsGenesRibo.daoy2d)
```

```{r removeMitochGenes}
# remove mitochondrial genes
filt.CellsGenesRiboMito.daoy2d <- filt.CellsGenesRibo.daoy2d[ ! grepl("^MT-", 
                                                                      rownames(filt.CellsGenesRibo.daoy2d)), ]

# check if ribosome and mitochondrial gene filtering works - it does:
# reduced from 24,628 genes to 24,615 genes
dim(filt.CellsGenesRibo.daoy2d)
dim(filt.CellsGenesRiboMito.daoy2d)
```

```
# save the cell, gene, ribosomal, mitochodrial gene filtered seurat object - no mitochondrial or 
# ribosomal genes.
saveRDS(filt.CellsGenesRiboMito.daoy2d, file = "./outputs50K/DAOY_2D/filt.CellsGenesRiboMito.daoy2dRep.seurat.rds")
```


--------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------
***EXPLORE SOURCES OF UNWANTED VARIATION, e.g. cell cycle***

Load the Seurat object  with only filtering of cells and genes
```{r load_filtered_CellsGenes_SeuratObject}
# the name of this object reflects it has undergone filtering for cells AND genes

fullFilt.daoy.2d.seurat <- readRDS("outputs50K/DAOY_2D/filtCellsGenes.daoy.2d.seurat.rds")
```

```{r normalize_fully_filtered_seuratObject}
# standard normalization function is sufficient to explore whether cells cluster by cell-cycle phase

seurat.phase <- NormalizeData(fullFilt.daoy.2d.seurat)
seurat.phase <- FindVariableFeatures(seurat.phase, selection.method = "vst")
seurat.phase <- ScaleData(seurat.phase, features = rownames(seurat.phase))
seurat.phase <- RunPCA(seurat.phase, features = VariableFeatures(seurat.phase))
```

```{r fig.height=6}
DimHeatmap(seurat.phase, dims = c(2,3,4))
```


```{r cellCycle_effects}
# Score cells for cell cycle.
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
seurat.phase <- CellCycleScoring(seurat.phase,
                                 g2m.features = g2m.genes,
                                 s.features = s.genes)
```

```{r}
head(seurat.phase@meta.data)
```

Looks like phase in the cell cycle has a significant effect on clustering!
```{r DimplotCellCycleEffect}
seurat.phase <- RunPCA(seurat.phase, features = c(s.genes, g2m.genes))
DimPlot(seurat.phase, reduction = "pca", group.by = "Phase")
```

Next step is to adjust for the cell cycle  (based on https://satijalab.org/seurat/articles/cell_cycle_vignette.html).
We do not want to entirely regress out the cell cycle, as we want to separate apart cycling from NON-cycling cells. However, differences in cell cycle phase among proliferating cells will be regressed out.
```{r }
seurat.phase$CC.difference <- seurat.phase$S.Score - seurat.phase$G2M.Score
seurat.phase <- ScaleData(seurat.phase, vars.to.regress = "CC.difference", features = rownames(seurat.phase))
```

```{r}
# cell cycle effects strongly mitigated in PCA
seurat.phase <- RunPCA(seurat.phase, features = VariableFeatures(seurat.phase), nfeatures.print = 10)
```
This worked! Compared to before regressing out the 'cell cycle difference', the DimPlot below recapitulates the final DimPlot on that webpage.
```
# when running a PCA on cell cycle genes, actively proliferating cells remain distinct from G1
# cells however, within actively proliferating cells, G2M and S phase cells group together.
# the two lines of code in this chunk are NECESSARY and should be run BEFORE the saveRDS() function!!!

seurat.phase <- RunPCA(seurat.phase, features = c(s.genes, g2m.genes))
DimPlot(seurat.phase,reduction = "pca", group.by = "Phase")
```

```
# save the seurat.phase object as a .RDS file in outputs50K
saveRDS(seurat.phase, file = "./outputs50K/DAOY_2D/seurat.phase.regressedCCdiff.daoy.2d.rds")
```

```{r}
seurat.phase.daoy2d <- readRDS("outputs50K/DAOY_2D/seurat.phase.regressedCCdiff.daoy.2d.rds")
```


```{r DimplotRegressOutCCdifference}
DimPlot(seurat.phase.daoy2d, reduction = "pca", group.by = "Phase")
```

The mitochondrial-based clustering shows no clustering by percent mitochondrial reads
```{r mitochondrial_expression}
# Check quartile values
summary(seurat.phase@meta.data$percent.mt)

# Turn percent.mt into categorical factor vector based on quartile values
seurat.phase@meta.data$mito.prop <- cut(seurat.phase@meta.data$percent.mt, 
                  breaks=c(-Inf, 1.8, 5.0, 9.99, Inf), 
                  labels=c("Low","Medium","Medium high", "High"))
```

```{r plot_mitochondrialExpression}
# repeat PCA and plot the PCA colored by mitochondrial percentage - this variable has no effect
# on the clustering.
seurat.phase <- RunPCA(seurat.phase)

DimPlot(seurat.phase,
        reduction = "pca",
        group.by= "mito.prop"
        )
```

-------------------------------------------------------------------------------------
***Using SCTransform with cell cycle regression***
```{r}
# normalise data
seurat.phase.sct <- NormalizeData(fullFilt.daoy.2d.seurat)

# run cell cycle scoring
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
seurat.phase.sct <- CellCycleScoring(seurat.phase.sct,
                                 g2m.features = g2m.genes,
                                 s.features = s.genes)

```

Normalisation, followed by CellCycleScoring, then SCTransform, then PCA worked! Get a similar result to DimplotRegressOutCCdifference code chunk above. Much faster to run. This code separates G1 from other cell cycle phases.
```{r DimplotSCTRegressOutCCdifference}
seurat.phase.sct$CC.difference <- seurat.phase.sct$S.Score - seurat.phase.sct$G2M.Score

seurat.phase.sct <- SCTransform(seurat.phase.sct,
                                   vars.to.regress = "CC.difference",
                                   method = "glmGamPoi", 
                                   verbose = FALSE)

seurat.phase.sct <- RunPCA(seurat.phase.sct, features = c(s.genes, g2m.genes))

DimPlot(seurat.phase.sct, reduction = "pca", group.by = "Phase")
```

```
# save the seurat.phase.sct object - this object still has ribosomal and mitochondrial genes

saveRDS(seurat.phase.sct, file = "./outputs50K/DAOY_2D/seurat.phase.SCT.regressedCCdiff.daoy.2d.rds")
```
-------------------------------------------------------------------------------------------
Run the SCTransform analysis and check whether the 'seurat.phase' objects look the same as the seurat object which has had mitochondrial and ribosomal genes removed. Answer - cell cycle dependent clustering is retained
```{r}
# load the seurat object
filt.CellsGenesRiboMito.daoy2d <- readRDS("outputs50K/DAOY_2D/filt.CellsGenesRiboMito.daoy2d.seurat.rds")

# normalise data
phase.daoy2d.filtRiboMito.sct <- NormalizeData(filt.CellsGenesRiboMito.daoy2d)

# run cell cycle scoring
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
phase.daoy2d.filtRiboMito.sct <- CellCycleScoring(phase.daoy2d.filtRiboMito.sct,
                                 g2m.features = g2m.genes,
                                 s.features = s.genes)

```

First run SCTransfrom WITHOUT vars.to.regress argument
```{r DimplotSCTRegressOutCCdifference}
# there is clustering by cell cycle
phase.daoy2d.filtRiboMito.sct <- SCTransform(phase.daoy2d.filtRiboMito.sct,
                                   
                                   method = "glmGamPoi", 
                                   verbose = FALSE)

phase.daoy2d.filtRiboMito.sct <- RunPCA(phase.daoy2d.filtRiboMito.sct, features = c(s.genes, g2m.genes))

DimPlot(phase.daoy2d.filtRiboMito.sct, reduction = "pca", group.by = "Phase")
```

Then run SCTransform with vars.to.regress argument = cell cycle difference
```{r DimplotSCTRegressOutCCdifferenceFiltRiboMito}
# now G1 and G2/M are just one cluster, separate from G1, i.e. there are 2 clusters instead of 3.

phase.daoy2d.filtRiboMito.sct$CC.difference <- phase.daoy2d.filtRiboMito.sct$S.Score - phase.daoy2d.filtRiboMito.sct$G2M.Score

phase.daoy2d.filtRiboMito.sct <- SCTransform(phase.daoy2d.filtRiboMito.sct,
                                   vars.to.regress = "CC.difference",
                                   method = "glmGamPoi", 
                                   verbose = FALSE)

phase.daoy2d.filtRiboMito.sct <- RunPCA(phase.daoy2d.filtRiboMito.sct, features = c(s.genes, g2m.genes))

DimPlot(phase.daoy2d.filtRiboMito.sct, reduction = "pca", group.by = "Phase")
```

```
# save the GENE, CELL, MITOCHONDRIAL and RIBOSOMAL FILTERED seurat object with regression of the cell cycle

saveRDS(phase.daoy2d.filtRiboMito.sct, file = "./outputs50K/DAOY_2D/phase.filtRiboMito.SCT.regressedCCdiff.daoy.2d.rds")
```









