---
title: "DAOY-SPHER_Seurat_EXPLORE1_50K"
author: "JJ"
date: "2023-01-26"
output: html_document
---

```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE,
                      fig.path = "./outputs50K/ONS76_SPHER/figures/")
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(glmGamPoi)
library(clustree)
library(tidyverse)
library(magrittr)
library(patchwork)
})
```

```{r}
# load the phase, cell cycle difference regressed, SCTransformed object

phase.daoySpher.filtRiboMito.sct <- readRDS("outputs50K/DAOY_SPHER/phase.daoySpher.filtRiboMito.SCT.regressedCCdiff.rds")
```

```{r}
# generate a clustered seurat object, res = 0.2
# PCA was previously run on 'phase.ons762d.filtRiboMito.sct' in the QC file

phase.daoySpher.sct.clust.res0.2 <- phase.daoySpher.filtRiboMito.sct %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, res = 0.2)
```


```{r}
# using res = 0.2 there are 2 clusters
DimPlot(phase.daoySpher.sct.clust.res0.2)
```

***Clustree***
```{r}
# attempt to use Clustree package to find the right number of clusters
# -----------------------------------------------------------------------------------------------

# Select a range of resolutions
resolution.range <- seq(from = 0, to = 1.8, by = 0.2)

# Find clusters using a range of resolutions, the clustering information is in the metadata slot
daoySpher.clusters <- FindClusters(object = phase.daoySpher.sct.clust.res0.2, 
                                      resolution = resolution.range,
                                      verbose = FALSE)
```

```{r ClustreeCBO, fig.height=6}
# the clustering tree finds 3 major clusters with res = 0.4.
# when res > 0.4 the cluster assignment becomes unstable.
# use marker differences (FindAllMarkers) and pathway enrichment to tell how many clusters have distinct enrichments. 
clustree(daoySpher.clusters)
```

Test how the clusters appear and check markers of each cluster
```{r}
# generate a clustered  seurat object, res = 0.4
phase.daoySpher.sct.clust.res0.4 <- phase.daoySpher.filtRiboMito.sct %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, res = 0.4)
```

```{r}
# using res = 0.4 there are 3 clusters, and only two of these are well separated
DimPlot(phase.daoySpher.sct.clust.res0.4)
```

```{r}
# get the number of cells per cluster
table(Idents(phase.daoySpher.sct.clust.res0.4))
```

***Find the DEGs of each cluster***
Find the markers of each cluster for res = 0.2
```{r}
# find POSITIVE markers of each cluster
phase.daoySpher.sct.POSdegs <- FindAllMarkers(object = phase.daoySpher.sct.clust.res0.2,
                                      min.pct = 0.25,
                                      only.pos = TRUE)
```

Cluster 0 gene expression is similar in terms of pct.1 and pct.2. However, cluster 1 gene expression shows marked differences from cluster 0
```{r}
phase.daoySpher.sct.POSdegs %>% 
  group_by(cluster) %>% 
  slice_head(n = 20)
```

Find the markers of each cluster for res = 0.4
```{r}
# find POSITIVE markers of each cluster
phase.daoySpher.sct.res0.4.POSdegs <- FindAllMarkers(object = phase.daoySpher.sct.clust.res0.4,
                                      min.pct = 0.25,
                                      only.pos = TRUE)
```

As can be seen below the markers for cluster 0 and cluster 1 are expressed by similar proportions of cells in pct.1 and pct.2, so RESOLTION = 0.2 IS APPROPRIATE
```{r}
phase.daoySpher.sct.res0.4.POSdegs %>% 
  group_by(cluster) %>% 
  slice_head(n = 20)
```

```{r}
# save the UP-regulated genes (res = 0.2)
saveRDS(phase.daoySpher.sct.POSdegs, 
        file = "./outputs50K/DAOY_SPHER/POS.DEGs.AllClusts.phase.daoySpher.res0.2.sct.rds")

```

```{r}
# load the markers object above
phase.daoySpher.sct.POSdegs <- 
  readRDS("outputs50K/DAOY_SPHER/POS.DEGs.AllClusts.phase.daoySpher.res0.2.sct.rds")
```

```{r}
# save the phase.daoySpher.sct.clust.res0.2 as a new Seurat object
saveRDS(phase.daoySpher.sct.clust.res0.2,
        file = "./outputs50K/DAOY_SPHER/phase.daoySpher.SCT.clustered.res0.2.rds")
```








