---
title: "DAOY-2Drep_Seurat_EXPLORE1_50K"
author: "JJ"
date: "2023-01-26"
output: html_document
---

```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE,
                      fig.path = "./outputs50K/ONS76_2D/figures/")
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(glmGamPoi)
library(presto)
library(clustree)
library(tidyverse)
library(magrittr)
library(patchwork)
})
```

```{r}
# load the data from scratch
# load the DAOY-2D raw data (from "working_data" folder)
data_dir <- "./working_data/DAOY-2D/"   # the dot is current dir and is essential!
list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r}
# note these are .gz files which can be directly read. n = 3838 cells
dyMono.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                         features = paste0(data_dir, "features.tsv.gz"), 
                         cells = paste0(data_dir, "barcodes.tsv.gz"))
# create the seurat object, keep genes expressed in 10 or more cells
dyMono <- CreateSeuratObject(counts = dyMono.data,
                          min.cells = 10, min.genes=2000,  # set high as this sample was very deeply sequenced
                          project = "dyMono")

dim(dyMono) # 24,957 genes and 2922 cells
```

```{r}
# identify mitochondrial genes
mito.genes <- grep(pattern = "^MT", x = rownames(x = dyMono), value = TRUE)
percent.mito <- 
  Matrix::colSums(dyMono[mito.genes, ])/Matrix::colSums(dyMono)

# identify ribosomal genes
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(x = dyMono), value = TRUE)
percent.ribo <- 
  Matrix::colSums(dyMono[ribo.genes, ])/Matrix::colSums(dyMono)

# add the ribosome and mitochondrial gene percent to the metadata
dyMono <- 
  AddMetaData(object = dyMono, metadata = percent.mito, col.name = "percent.mito")

dyMono <- 
  AddMetaData(object = dyMono, metadata = percent.ribo, col.name = "percent.ribo")

# remove ribosomal and mitochondrial genes
# remove ribosomal genes first
dyMono <- dyMono[ ! grepl('^RP[SL]', rownames(dyMono)), ]

# remove mitochondrial genes 
dyMono <- dyMono[ ! grepl("^MT-", rownames(dyMono)), ]

dim(dyMono) # 24,846 genes and 2922 cells
```

```{r}
# remove all cells with percent.mito > 10%
dyMono <- subset(dyMono, subset = percent.mito < 0.1)
dim(dyMono) # reduced to 15,846 genes and 2633 cells
```

```{r}
# save the NON-normalised, NON-clustered seurat object
saveRDS(dyMono, file = "./outputs50K/DAOY_2D/filt.RiboMitoCellsGenes.seurat.DaoyMonolayer.rds")
```

```{r}
dyMono <- readRDS("outputs50K/DAOY_2D/filt.RiboMitoCellsGenes.seurat.DaoyMonolayer.rds")
```

```{r}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference

# normalise object before running cell cycle scores as per Seurat github issue: https://github.com/satijalab/seurat/issues/1679
dyMono <- NormalizeData(dyMono)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
dyMono <- CellCycleScoring(dyMono,
                g2m.features = g2m.genes,
                s.features = s.genes)
dyMono$CC.difference <- dyMono$S.Score - dyMono$G2M.Score
```

```{r}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
dyMono.input.integration <- SCTransform(dyMono, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```{r}
# save the SCTransformed object, CC difference regressed, PCA reduction, as input for integrated workflow
saveRDS(dyMono.input.integration, "./outputs50K/DAOY_2D/DAOYmono.inputIntegration.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```

```{r}
dyMono <- SCTransform(dyMono, vars.to.regress = "CC.difference", method = "glmGamPoi", verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.5)
```

```{r}
# save the clustered, SCTransformed object above
saveRDS(dyMono, file = "./outputs50K/DAOY_2D/DAOYmono.RiboMitoCellsGenesFilt.SCT.clustered.CCdiffRegressed.rds")
```

```{r}
# load the above object.
dyMono <- readRDS("outputs50K/DAOY_2D/DAOYmono.RiboMitoCellsGenesFilt.SCT.clustered.CCdiffRegressed.rds")
```

```{r DimplotDAOYmonoRepUmapSCTransform, fig.height=6, fig.width=10}
DimPlot(dyMono, label = TRUE)
```

```{r}
# get the markers of each cluster - slow to run
dyMono.POSmarkers <- 
  FindAllMarkers(object = dyMono, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
dyMono.POSmarkers %>% 
  dplyr::group_by(cluster) %>% 
  slice_head(n = 30)
```

```{r}
# None of these five genes is expressed
dyMono.POSmarkers %>%
  filter(gene %in% c("NEUROD1", "CNTN1", "CNTN2", "MEIS1", "DCX1"))
```

```{r}
# save the positive markers of DAOY-2D (rep)
saveRDS(dyMono.POSmarkers, 
        file = "./outputs50K/DAOY_2D/POS.markers.DAOYmonolayerRep.SCT.clustered.rds")
```

```{r}
dyMono.POSmarkers <- readRDS("outputs50K/DAOY_2D/POS.markers.DAOYmonolayerRep.SCT.clustered.rds")
```

```{r HeatmapDaoyMonoTop10markers5clustersRes0.5, fig.height=6, fig.width=6}
# first load dyMono seurat object above, then only run this chunk
dyMono.POSmarkers %>% 
  group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(dyMono, features = top10$gene) + NoLegend()
```

As can be seen from the above heatmap the res=0.5 clustering seems to generate too many clusters. For example, cluster 2 has many genes in common with cluster 1. Also cluster 3 and cluster 5 share many markers in common. Try resolution=0.2 and make anew heatmap

```{r}
# first load "filt.RiboMitoCellsGenes.seurat.DaoyMonolayer.rds" then normalise, run CC regression,
# then cluster the cells as per this code chunk
dyMono.res0.2 <- SCTransform(dyMono, vars.to.regress = "CC.difference", method = "glmGamPoi", verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.2) # for res=0.2, clustree shows there are 4 clusters
```

```{r}
# get the markers of each cluster - slow to run
dyMono.POSmarkers.res0.2 <- 
  FindAllMarkers(object = dyMono.res0.2, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# save the POS markers at res=0.2
saveRDS(dyMono.POSmarkers.res0.2,
        file = "./outputs50K/DAOY_2D/POS.markers.res0.2.DAOYmonoRep.SCT.clustered.rds")
```


```{r HeatmapDaoyMonoTop10markers4clustersRes0.2, fig.height=5, fig.width=6}
dyMono.POSmarkers.res0.2 %>% 
  group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10.res.0.2
DoHeatmap(dyMono.res0.2, features = top10.res.0.2$gene) + guides(color = "none")
```

The above clusters are better defined. Cluster 0, 1, and 3 are well separated, and cluster 2 has lower level expression of genes that are also expressed in cluster 0 and 1, but not in cluster 3. So run presto, then gProfiler on this sample to get an idea of what pathways are active in the different cell clusters.

```{r}
saveRDS(dyMono.res0.2, 
        file = "./outputs50K/DAOY_2D/DAOYmono.RiboMitoCellsGenesFilt.SCT.res0.2.clustered.CCdiffRegressed.rds")
```

```{r}
# prepare the data for gProfiler (web app) input
dyMono.res0.2.genes.presto <- wilcoxauc(dyMono.res0.2, 'seurat_clusters')
nrow(dyMono.res0.2.genes.presto)
dplyr::count(dyMono.res0.2.genes.presto, group)
```

```{r}
head(dyMono.res0.2.genes.presto, 3)
```

```{r}
# filter and order the gene list
dyMono.res0.2.genes.presto.clust0 <- dyMono.res0.2.genes.presto %>% 
  dplyr::filter(group == "0" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.res0.2.genes.presto.clust0) # 1331 genes
```

```{r}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(dyMono.res0.2.genes.presto.clust0, 
          "./outputs50K/DAOY_2D/daoyMono.res0.2.cluster0.genes.presto.gProfilerInput.csv", quote = F)
```

```{r}
# filter and order the gene list
dyMono.res0.2.genes.presto.clust1 <- dyMono.res0.2.genes.presto %>% 
  dplyr::filter(group == "1" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.res0.2.genes.presto.clust1) # 699 genes
```

```{r}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(dyMono.res0.2.genes.presto.clust1, 
          "./outputs50K/DAOY_2D/daoyMono.res0.2.cluster1.genes.presto.gProfilerInput.csv", quote = F)
```

```{r}
# filter and order the gene list
# AS THERE ARE ONLY 2 MARKERS, DO NOT SAVE THE LIST, AND DO NOT RUN gPROFILER FOR THIS CLUSTER
dyMono.res0.2.genes.presto.clust2 <- dyMono.res0.2.genes.presto %>% 
  dplyr::filter(group == "2" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.res0.2.genes.presto.clust2) # 2 genes only! The reason for this is apparent on looking at the heatmap.
```

```{r}
# filter and order the gene list
dyMono.res0.2.genes.presto.clust3 <- dyMono.res0.2.genes.presto %>% 
  dplyr::filter(group == "3" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(dyMono.res0.2.genes.presto.clust3) # 1142 genes
```

```{r}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(dyMono.res0.2.genes.presto.clust3, 
          "./outputs50K/DAOY_2D/daoyMono.res0.2.cluster3.genes.presto.gProfilerInput.csv", quote = F)
```




***Clustree***
```{r}
# attempt to use Clustree package to find the right number of clusters
# -----------------------------------------------------------------------------------------------

# Select a range of resolutions
resolution.range <- seq(from = 0, to = 1.8, by = 0.2)

# Find clusters using a range of resolutions, the clustering information is in the metadata slot
dyMono.clusters <- FindClusters(object = dyMono, 
                                      resolution = resolution.range,
                                      verbose = FALSE)
```

```{r ClustreeCBO, fig.height=8}
# the clustering tree finds too many clusters.
# with res = 0.8 there are 4 clusters; with res = 1 there are 5 clusters.
# use marker differences (FindAllMarkers) and pathway enrichment to tell how many clusters have distinct enrichments. 

clustree(dyMono.clusters)
```

```{r}
# get the number of cells per cluster
table(Idents(dyMono.clusters))
```







