---
title: "hovestadt_nature_2019_smartSeq2_QC_and_analysis"
author: "JJ"
date: "2023-03-02"
output: html_document
---

```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE,
                      fig.path = "./outputs50K/Hovestadt_outputs/figures/")
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(SeuratData)
library(SeuratWrappers)
library(batchelor)
library(cluster)
library(factoextra)
library(glmGamPoi)
library(Matrix)
library(patchwork)
library(tidyverse)
})
```


***Cluster MUV41 patient SHH-MB***
```{r eval=FALSE, echo=FALSE}
raw.counts.muv41 <- read.table(file = "./hovestadt_Nature2019/GSM3905417_MUV41.txt",
                         sep = "")
```

```{r eval=FALSE, echo=FALSE}
head(raw.counts.muv41)
```

```{r eval=FALSE, echo=FALSE}
muv41 <- CreateSeuratObject(raw.counts.muv41,
                            min.cells = 10,
                            min.genes = 2500,
                            project = "muv41")
```

```{r eval=FALSE, echo=FALSE}
dim(muv41)
```

```{r eval=FALSE, echo=FALSE}
muv41 <- SCTransform(muv41, vst.flavor = "v2", verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE) %>% 
  RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindNeighbors(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindClusters(resolution = 0.7, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(muv41,
        file = "./outputs50K/Hovestadt_outputs/muv41.PatientSHHMB.SCT.clustered.rds")
```

```{r}
muv41 <- readRDS("outputs50K/Hovestadt_outputs/muv41.PatientSHHMB.SCT.clustered.rds")
```


```{r eval=FALSE, echo=FALSE}
DimPlot(muv41, label = T, repel = T)
```

```{r ViolinPlotMUV41shhMBneurod1Rtn1Rbfox3, fig.height=4, fig.width=12}
VlnPlot(muv41, features = c("NEUROD1", "RBFOX3", "RTN1"))
```

```{r eval=FALSE, echo=FALSE}
# the aim is to get cell cycle phase for all cells in the below code chunks
muv41.unclust <- CreateSeuratObject(raw.counts.muv41,
                            min.cells = 10,
                            min.genes = 2500,
                            project = "muv41")
```

```{r eval=FALSE, echo=FALSE}
saveRDS(muv41.unclust,
        file = "./outputs50K/Hovestadt_outputs/MUV41.unclustered.PatientSHHMB.rds")
```

```{r eval=FALSE, echo=FALSE}
# get cell cycle info
muv41.unclust.norm <- NormalizeData(muv41.unclust)
# include normalisation step of dyCbo before cell cycle scoring as per github issue: https://github.com/satijalab/seurat/issues/1679
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
muv41.unclust.norm <- CellCycleScoring(muv41.unclust.norm,
                g2m.features = g2m.genes,
                s.features = s.genes)

muv41.phase <- SCTransform(muv41.unclust.norm, vst.flavor = "v2", verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE) %>% 
  RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindNeighbors(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindClusters(resolution = 0.7, verbose = FALSE)
```
****Isolate NEUROD1+ cells for MUV41****
```{r eval=FALSE, echo=FALSE}
# isolate all NEUROD1 expressing cells, there are 303 cells
neurod1.muv41 <- GetAssayData(object = muv41.phase,
                            assay = "RNA",      # choosing "SCT" assay makes very little difference!
                            slot = "data")["NEUROD1", ]
pos.ids.neurod1.muv41 <- names(which(neurod1.muv41 > 0))
length(pos.ids.neurod1.muv41)
```
```{r eval=FALSE, echo=FALSE}
neurod1.pos.muv41.phase <- subset(muv41.phase, cells = pos.ids.neurod1.muv41)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(neurod1.pos.muv41.phase,
        file = "./outputs50K/Hovestadt_outputs/NeuroD1positiveCells.MUV41.PatientSHH-MB.rds")
```


```{r eval=FALSE, echo=FALSE}
# transfer meta.data to a new object called metadata.dy
metadata.neurod1.muv41.phase <- neurod1.pos.muv41.phase@meta.data %>% 
  rownames_to_column(var = "cellID") # this step needed as when 'select' is applied later on, the rownames
                                    # (cell names) are still retained
```

```{r eval=FALSE, echo=FALSE}
# subset the metadata dataframe to keep just 'orig.ident' and 'Phase' columns (cc = cell cycle phase)
metadata.cc.neurod1.muv41.phase <- metadata.neurod1.muv41.phase %>% 
  dplyr::select(orig.ident, Phase)
# for each orig.ident get a count for each orig.ident and for each Phase
summ.stats.cc.neurod1.muv41.phase <- metadata.cc.neurod1.muv41.phase %>% 
  dplyr::group_by(Phase, orig.ident) %>% 
  dplyr::summarise(count= n())

summ.stats.cc.neurod1.muv41.phase  # total is 303 cells. G1=179, G2M=48, S=76
```

```{r eval=FALSE, echo=FALSE}
dim(metadata.cc.neurod1Clus6cbo)
```

```{r eval=FALSE, echo=FALSE}
# convert the counts to a proportion where G1 count is divided by the sum of G1 + G2M + S counts
summ.stats.props.cc.neurod1.muv41.phase <- summ.stats.cc.neurod1.muv41.phase %>% 
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(props = count / sum(count))

summ.stats.props.cc.neurod1.muv41.phase
```


***SJ454 SHH-MB***
```{r eval=FALSE, echo=FALSE}
raw.counts.sj454 <- read.table(file = "./hovestadt_Nature2019/GSM3905423_SJ454.txt",
                         sep = "")
```

```{r eval=FALSE, echo=FALSE}
head(raw.counts.sj454)
```

```{r eval=FALSE, echo=FALSE}
sj454 <- CreateSeuratObject(raw.counts.sj454,
                            min.cells = 10,
                            min.genes = 2500,
                            project = "reanalysis_hovestadt")
```

```{r eval=FALSE, echo=FALSE}
dim(sj454)
```

```{r eval=FALSE, echo=FALSE}
sj454 <- SCTransform(sj454, vst.flavor = "v2", verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE) %>% 
  RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindNeighbors(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindClusters(resolution = 0.7, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
DimPlot(sj454, label = T, repel = T)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(sj454,
        file = "./outputs50K/Hovestadt_outputs/SJ454.SHH_MB.SOX2networkExpression.MitoCellsGenesFilt.RiboRetained.SCT.clustered.rds")
```

```{r}
sj454.clust <- readRDS("outputs50K/Hovestadt_outputs/SJ454.SHH_MB.SOX2networkExpression.MitoCellsGenesFilt.RiboRetained.SCT.clustered.rds")
```

```{r ViolinPlotSJ454shhMBneurod1Rtn1Rbfox3, fig.height=4, fig.width=12}
VlnPlot(sj454.clust, features = c("NEUROD1", "RBFOX3", "RTN1"))
```

****Isolate NEUROD1+ cells for SJ454****
```{r eval=FALSE, echo=FALSE}
# the aim is to get cell cycle phase for all cells in the below code chunks
sj454.unclust <- CreateSeuratObject(raw.counts.sj454,
                            min.cells = 10,
                            min.genes = 2500,
                            project = "sj454")
```

```{r eval=FALSE, echo=FALSE}
saveRDS(sj454.unclust,
        file = "./outputs50K/Hovestadt_outputs/SJ454.unclustered.PatientSHHMB.rds")
```

```{r eval=FALSE, echo=FALSE}
# get cell cycle info
sj454.unclust.norm <- NormalizeData(sj454.unclust)
# include normalisation step of dyCbo before cell cycle scoring as per github issue: https://github.com/satijalab/seurat/issues/1679
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
sj454.unclust.norm <- CellCycleScoring(sj454.unclust.norm,
                g2m.features = g2m.genes,
                s.features = s.genes)

sj454.phase <- SCTransform(sj454.unclust.norm, vst.flavor = "v2", verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE) %>% 
  RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindNeighbors(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindClusters(resolution = 0.7, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
# isolate all NEUROD1 expressing cells, there are 303 cells
neurod1.sj454 <- GetAssayData(object = sj454.phase,
                            assay = "RNA",      # choosing "SCT" assay makes very little difference!
                            slot = "data")["NEUROD1", ]
pos.ids.neurod1.sj454 <- names(which(neurod1.sj454 > 0))
length(pos.ids.neurod1.sj454)
```
```{r eval=FALSE, echo=FALSE}
neurod1.pos.sj454.phase <- subset(sj454.phase, cells = pos.ids.neurod1.sj454)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(neurod1.pos.sj454.phase,
        file = "./outputs50K/Hovestadt_outputs/NeuroD1positiveCells.SJ454.PatientSHH-MB.rds")
```


```{r eval=FALSE, echo=FALSE}
# transfer meta.data to a new object called metadata.dy
metadata.neurod1.sj454.phase <- neurod1.pos.sj454.phase@meta.data %>% 
  rownames_to_column(var = "cellID") # this step needed as when 'select' is applied later on, the rownames
                                    # (cell names) are still retained
```

```{r eval=FALSE, echo=FALSE}
# subset the metadata dataframe to keep just 'orig.ident' and 'Phase' columns (cc = cell cycle phase)
metadata.cc.neurod1.sj454.phase <- metadata.neurod1.sj454.phase %>% 
  dplyr::select(orig.ident, Phase)
# for each orig.ident get a count for each orig.ident and for each Phase
summ.stats.cc.neurod1.sj454.phase <- metadata.cc.neurod1.sj454.phase %>% 
  dplyr::group_by(Phase, orig.ident) %>% 
  dplyr::summarise(count= n())

summ.stats.cc.neurod1.sj454.phase  # total is 303 cells. G1=179, G2M=48, S=76
```

```{r eval=FALSE, echo=FALSE}
# convert the counts to a proportion where G1 count is divided by the sum of G1 + G2M + S counts
summ.stats.props.cc.neurod1.sj454.phase <- summ.stats.cc.neurod1.sj454.phase %>% 
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(props = count / sum(count))

summ.stats.props.cc.neurod1.sj454.phase. # total=238 cells. G1=0.69, G2M= 0.09, S=0.21
```



```{r eval=FALSE, echo=FALSE, VlnplotSOX2regNetworkSRRTgeneSJ454shhMB, fig.height=14, fig.width=22}
# cluster 2 expresses all NSC factors except OTX2
# VlnPlot(sj454, features = c("SOX2", "POU3F2", "FABP7", "OTX2", "ZIC2", "GLI2", "OCT4", "PAX6"))

 p1a <- VlnPlot(sj454.clust, 
        features = c("FABP7")) +
 theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1b <- VlnPlot(sj454.clust, 
        features = c("ZIC2")) +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1c <- VlnPlot(sj454.clust, 
        features = c("SOX2")) +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1d <- VlnPlot(sj454.clust, 
        features = c("SRRT")) +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1e <- VlnPlot(sj454.clust, 
        features = c("POU3F2")) +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1f <- VlnPlot(sj454.clust, 
        features = c("GLI2")) +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1a + p1b + p1c + p1d + p1e + p1f + plot_layout(ncol = 2)
```


***SJ577 SHH-MB***
```{r eval=FALSE, echo=FALSE}
raw.counts.sj577 <- read.table(file = "./hovestadt_Nature2019/GSM3905425_SJ577.txt",
                         sep = "")
```

```{r eval=FALSE, echo=FALSE}
head(raw.counts.sj577)
```

```{r eval=FALSE, echo=FALSE}
sj577 <- CreateSeuratObject(raw.counts.sj577,
                            min.cells = 10,
                            min.genes = 2500,
                            project = "reanalysis_hovestadt")
```

```{r eval=FALSE, echo=FALSE}
dim(sj577)
```

```{r eval=FALSE, echo=FALSE}
sj577 <- SCTransform(sj577, vst.flavor = "v2", verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE) %>% 
  RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindNeighbors(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindClusters(resolution = 0.7, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(sj577,
        file = "./outputs50K/Hovestadt_outputs/SJ577.PatientSHH_MB.SCT.clustered.rds")
```

```{r}
sj577 <- readRDS("outputs50K/Hovestadt_outputs/SJ577.PatientSHH_MB.SCT.clustered.rds")
```


```{r eval=FALSE, echo=FALSE}
DimPlot(sj577, label = T, repel = T)
```

```{r ViolinPlotSJ577shhMBneurod1Rtn1Rbfox3, fig.height=4, fig.width=12}
VlnPlot(sj577, features = c("NEUROD1", "RBFOX3", "RTN1"))
```


****Isolate NEUROD1+ cells from SJ577****
```{r eval=FALSE, echo=FALSE}
raw.counts.sj577 <- read.table(file = "./hovestadt_Nature2019/GSM3905425_SJ577.txt",
                         sep = "")
```

```{r eval=FALSE, echo=FALSE}
# the aim is to get cell cycle phase for all cells in the below code chunks
sj577.unclust <- CreateSeuratObject(raw.counts.sj577,
                            min.cells = 10,
                            min.genes = 2500,
                            project = "sj577")
```

```{r eval=FALSE, echo=FALSE}
saveRDS(sj577.unclust,
        file = "./outputs50K/Hovestadt_outputs/SJ577.unclustered.PatientSHHMB.rds")
```


```{r eval=FALSE, echo=FALSE}
# get cell cycle info
sj577.unclust.norm <- NormalizeData(sj577.unclust)
# include normalisation step of dyCbo before cell cycle scoring as per github issue: https://github.com/satijalab/seurat/issues/1679
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
sj577.unclust.norm <- CellCycleScoring(sj577.unclust.norm,
                g2m.features = g2m.genes,
                s.features = s.genes)

sj577.phase <- SCTransform(sj577.unclust.norm, vst.flavor = "v2", verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE) %>% 
  RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindNeighbors(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
  FindClusters(resolution = 0.7, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
# isolate all NEUROD1 expressing cells, there are 303 cells
neurod1.sj577 <- GetAssayData(object = sj577.phase,
                            assay = "RNA",      # choosing "SCT" assay makes very little difference!
                            slot = "data")["NEUROD1", ]
pos.ids.neurod1.sj577 <- names(which(neurod1.sj577 > 0))
length(pos.ids.neurod1.sj577)
```
```{r eval=FALSE, echo=FALSE}
neurod1.pos.sj577.phase <- subset(sj577.phase, cells = pos.ids.neurod1.sj577)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(neurod1.pos.sj577.phase,
        file = "./outputs50K/Hovestadt_outputs/NeuroD1positiveCells.SJ577.PatientSHH-MB.rds")
```


```{r eval=FALSE, echo=FALSE}
# transfer meta.data to a new object called metadata.dy
metadata.neurod1.sj577.phase <- neurod1.pos.sj577.phase@meta.data %>% 
  rownames_to_column(var = "cellID") # this step needed as when 'select' is applied later on, the rownames
                                    # (cell names) are still retained
```

```{r eval=FALSE, echo=FALSE}
# subset the metadata dataframe to keep just 'orig.ident' and 'Phase' columns (cc = cell cycle phase)
metadata.cc.neurod1.sj577.phase <- metadata.neurod1.sj577.phase %>% 
  dplyr::select(orig.ident, Phase)
# for each orig.ident get a count for each orig.ident and for each Phase
summ.stats.cc.neurod1.sj577.phase <- metadata.cc.neurod1.sj577.phase %>% 
  dplyr::group_by(Phase, orig.ident) %>% 
  dplyr::summarise(count= n())

summ.stats.cc.neurod1.sj577.phase  # total is 257 cells. G1=166, G2M=33, S=58
```

```{r eval=FALSE, echo=FALSE}
# convert the counts to a proportion where G1 count is divided by the sum of G1 + G2M + S counts
summ.stats.props.cc.neurod1.sj577.phase <- summ.stats.cc.neurod1.sj577.phase %>% 
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(props = count / sum(count))

summ.stats.props.cc.neurod1.sj577.phase # total=238 cells. G1=0.69, G2M= 0.09, S=0.21
```


------------------------------------------------------------------------------------------------------
***Prepare DAOY-only cells in coculture for a merge with other datasets***
```{r eval=FALSE, echo=FALSE}
# load the DAOY-CBO object
dyCbo <- readRDS("outputs50K/DAOY-CBO/filt.MitoCellsGenes.RiboGenesRetained.seurat.DAOYcbo.rds")
```

```{r eval=FALSE, echo=FALSE}
dyOnly.dyCbo <- 
  readRDS("outputs50K/DAOY-CBO/DAOYOnlyCells.SubsetDAOYCboSample.Annotated.Rescaled.SCT.NoRiboNoMito.CCdiffRegressed.rds")
```

```{r eval=FALSE, echo=FALSE}
# get the DAOY only cells that have NOT been through SCT normalisation
dyOnlyIDs <- WhichCells(dyOnly.dyCbo)
dyOnly <- subset(dyCbo, cells = dyOnlyIDs)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(dyOnly,
        "./outputs50K/INTEGRATE/DAOYonly.DAOYcbo.NOTclustered.NoSCT.FiltMitoCellsGenes.RiboRetained.rds")
```

```{r eval=FALSE, echo=FALSE}
dyOnly <- 
  readRDS("outputs50K/INTEGRATE/DAOYonly.DAOYcbo.NOTclustered.NoSCT.FiltMitoCellsGenes.RiboRetained.rds")
```



------------------------------------------------------------------------------------------------------
***Integrate SHH-MB, DAOY-only coculture using fastMNN***

```{r eval=FALSE, echo=FALSE}
# identify the files of interest
list.files('hovestadt_Nature2019')
```

```{r eval=FALSE, echo=FALSE}
# load all the datasets to be merged in different combinations
# 3 SHH-MB
raw.counts.muv41 <- read.table(file = "./hovestadt_Nature2019/GSM3905417_MUV41.txt",
                         sep = "")
raw.counts.sj454 <- read.table(file = "./hovestadt_Nature2019/GSM3905423_SJ454.txt",
                         sep = "")
raw.counts.sj577 <- read.table(file = "./hovestadt_Nature2019/GSM3905425_SJ577.txt",
                         sep = "")

# 2 PDX SHH-MB
raw.counts.rmcb18 <- read.table(file = "./hovestadt_Nature2019/GSM3905439_RCMB18.txt",
                         sep = "")
raw.counts.rmcb24 <- read.table(file = "./hovestadt_Nature2019/GSM3905441_RCMB24.txt",
                         sep = "")

# load the DAOY spheroid data
dySpher <- readRDS("outputs50K/DAOY_SPHER/filt.MitoCellsGenes.RiboRetained.seurat.DAOYspher.rds")

# load the DAOY-only coculture data (not been through SCT normalisation or clustering)
dyOnly <- 
  readRDS("outputs50K/INTEGRATE/DAOYonly.DAOYcbo.NOTclustered.NoSCT.FiltMitoCellsGenes.RiboRetained.rds")


# create Seurat objects for tumours - 3 patient SHH-MB and 2 PDX
muv41.shh <- CreateSeuratObject(raw.counts.muv41,
                            min.cells = 10,
                            min.genes = 200,
                            project = "SHH-MB1")
sj454.shh <- CreateSeuratObject(raw.counts.sj454,
                            min.cells = 10,
                            min.genes = 200,
                            project = "SHH-MB2")
sj577.shh <- CreateSeuratObject(raw.counts.sj577,
                            min.cells = 10,
                            min.genes = 200,
                            project = "SHH-MB3")

rmcb18.pdx <- CreateSeuratObject(raw.counts.rmcb18,
                            min.cells = 10,
                            min.genes = 200,
                            project = "PDX1")
rmcb24.pdx <- CreateSeuratObject(raw.counts.rmcb24,
                            min.cells = 10,
                            min.genes = 200,
                            project = "PDX2")
```

```{r eval=FALSE, echo=FALSE}
# merge 3 patient SHH-MB tumours with DAOY-only cells in DAOY-CBO coculture
hov.3shhMbtumours.dyOnly.merged <- merge(muv41.shh, 
                    y = c(sj454.shh, sj577.shh, dyOnly),
                    project = "hov.3shhMbtumours.dyOnly.merged")
```

```{r eval=FALSE, echo=FALSE}
metadata <- hov.3shhMbtumours.dyOnly.merged@meta.data
# change orig.ident as below
metadata$orig.ident[which(str_detect(metadata$orig.ident, "dyCbo"))] <- "DAOY_coculture"
```

```{r eval=FALSE, echo=FALSE}
hov.3shhMbtumours.dyOnly.merged@meta.data <- metadata
```

```{r eval=FALSE, echo=FALSE}
table(hov.3shhMbtumours.dyOnly.merged@meta.data$orig.ident)
```

```{r eval=FALSE, echo=FALSE}
# run fastMNN on DAOY-only cells in coculture and patient SHH-MB
# scale data step is not necessary as RunFastMNN does its own scaling
hov.3shhMbtumours.dyOnly.merged.clust <- NormalizeData(hov.3shhMbtumours.dyOnly.merged)
hov.3shhMbtumours.dyOnly.merged.clust <- FindVariableFeatures(hov.3shhMbtumours.dyOnly.merged.clust)
hov.3shhMbtumours.dyOnly.merged.clust <- RunFastMNN(object.list = SplitObject(hov.3shhMbtumours.dyOnly.merged.clust, split.by = "orig.ident"))
hov.3shhMbtumours.dyOnly.merged.clust <- RunUMAP(hov.3shhMbtumours.dyOnly.merged.clust, reduction = "mnn", dims = 1:20)
hov.3shhMbtumours.dyOnly.merged.clust <- FindNeighbors(hov.3shhMbtumours.dyOnly.merged.clust, reduction = "mnn", dims = 1:20)
hov.3shhMbtumours.dyOnly.merged.clust <- FindClusters(hov.3shhMbtumours.dyOnly.merged.clust)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(hov.3shhMbtumours.dyOnly.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/DAOYonlyDAOYcbo.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
hov.3shhMbtumours.dyOnly.merged.clust <- 
  readRDS("outputs50K/Hovestadt_outputs/DAOYonlyDAOYcbo.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotHOVSHHMBtumoursDaoyOnlyCocultMNN, fig.height=6, fig.width=6}
p2 <- DimPlot(hov.3shhMbtumours.dyOnly.merged.clust, group.by = c("orig.ident"), pt.size = 3) +
  theme(legend.text = element_text(size = 20),
        plot.title = element_blank())
p2
```


***Run fastMNN on DAOY-spheroid in comparison to SHH-MB's only before calculating Mixing Metric***

```{r eval=FALSE, echo=FALSE}
# merge daoy-spher with patient SHH-MB's only
dySpher.hovShhMB.merged <- merge(dySpher,
                                 y = c(muv41.shh, sj454.shh, sj577.shh),
                                 project = "hov.SHHMB.dySpher.merged")
```

```{r eval=FALSE, echo=FALSE}
metadata <- dySpher.hovShhMB.merged@meta.data
# change orig.ident as below
metadata$orig.ident[which(str_detect(metadata$orig.ident, "dySpher"))] <- "DAOY_spheroid"
```

```{r eval=FALSE, echo=FALSE}
dySpher.hovShhMB.merged@meta.data <- metadata
```

```{r eval=FALSE, echo=FALSE}
# scale data step is not necessary as RunFastMNN does its own scaling
dySpher.hovShhMB.merged.clust <- NormalizeData(dySpher.hovShhMB.merged)
dySpher.hovShhMB.merged.clust <- FindVariableFeatures(dySpher.hovShhMB.merged.clust)
dySpher.hovShhMB.merged.clust <- RunFastMNN(object.list = SplitObject(dySpher.hovShhMB.merged.clust, split.by = "orig.ident"))
dySpher.hovShhMB.merged.clust <- RunUMAP(dySpher.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
dySpher.hovShhMB.merged.clust <- FindNeighbors(dySpher.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
dySpher.hovShhMB.merged.clust <- FindClusters(dySpher.hovShhMB.merged.clust)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(dySpher.hovShhMB.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/DAOYspher.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
dySpher.hovShhMB.merged.clust <- 
  readRDS("outputs50K/Hovestadt_outputs/DAOYspher.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotHOVSHHMBtumoursDAOYspherMNN, fig.height=6, fig.width=6}
p3 <- DimPlot(dySpher.hovShhMB.merged.clust, group.by = c("orig.ident"), pt.size = 3) +
  theme(legend.text = element_text(size = 20),
        plot.title = element_blank())
p3
```

***Run fastMNN on patient SHH-MB's only***

```{r eval=FALSE, echo=FALSE}
# only merge the three SHH-MB's
hovShhMB.merged <- merge(muv41.shh,
                        y = c(sj454.shh, sj577.shh),
                        project = "hov.SHHMB.merged")
```

```{r eval=FALSE, echo=FALSE}
# scale data step is not necessary as RunFastMNN does its own scaling
hovShhMB.merged.clust <- NormalizeData(hovShhMB.merged)
hovShhMB.merged.clust <- FindVariableFeatures(hovShhMB.merged.clust)
hovShhMB.merged.clust <- RunFastMNN(object.list = SplitObject(hovShhMB.merged.clust, split.by = "orig.ident"))
hovShhMB.merged.clust <- RunUMAP(hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
hovShhMB.merged.clust <- FindNeighbors(hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
hovShhMB.merged.clust <- FindClusters(hovShhMB.merged.clust)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(hovShhMB.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
hovShhMB.merged.clust <- 
  readRDS("outputs50K/Hovestadt_outputs/SHHMBHov.fastMNN.clust.rds")
```


```{r eval=FALSE, echo=FALSE, DimplotHOVSHHMBtumoursMNN, fig.height=6, fig.width=6}
p4 <- DimPlot(hovShhMB.merged.clust, group.by = c("orig.ident"), pt.size = 3) +
  theme(legend.text = element_text(size = 20),
        plot.title = element_blank())
p4
```


***Run fastMNN integration on patient SHH-MB's and PDX-RMCB18***
```{r eval=FALSE, echo=FALSE}
# merge RMCB18-PDX first with patient SHH-MBs
rmcb18.hovShhMB.merged <- merge(muv41.shh,
                                 y = c(sj454.shh, sj577.shh, rmcb18.pdx),
                                 project = "hov.SHHMB.rmcb18.merged")
```

```{r eval=FALSE, echo=FALSE}
# scale data step is not necessary as RunFastMNN does its own scaling
rmcb18.hovShhMB.merged.clust <- NormalizeData(rmcb18.hovShhMB.merged)
rmcb18.hovShhMB.merged.clust <- FindVariableFeatures(rmcb18.hovShhMB.merged.clust)
rmcb18.hovShhMB.merged.clust <- RunFastMNN(object.list = SplitObject(rmcb18.hovShhMB.merged.clust, split.by = "orig.ident"))
rmcb18.hovShhMB.merged.clust <- RunUMAP(rmcb18.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
rmcb18.hovShhMB.merged.clust <- FindNeighbors(rmcb18.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
rmcb18.hovShhMB.merged.clust <- FindClusters(rmcb18.hovShhMB.merged.clust)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(rmcb18.hovShhMB.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/RMCB18PDX.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
rmcb18.hovShhMB.merged.clust <- 
  readRDS("outputs50K/Hovestadt_outputs/RMCB18PDX.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotHOVSHHMBtumoursRMCB18PDXfastMNN, fig.height=6, fig.width=6}
p5 <- DimPlot(rmcb18.hovShhMB.merged.clust, group.by = c("orig.ident"), pt.size = 3) +
  theme(legend.text = element_text(size = 20),
        plot.title = element_blank())
p5
```


```{r eval=FALSE, echo=FALSE}
DimPlot(rmcb18.hovShhMB.merged.clust,
        "tsne") 
```


***Run fastMNN integration on patient SHH-MB's and PDX-RMCB24***
```{r eval=FALSE, echo=FALSE}
# merge RMCB24-PDX with patient SHH-MBs
rmcb24.hovShhMB.merged <- merge(muv41.shh,
                                 y = c(sj454.shh, sj577.shh, rmcb24.pdx),
                                 project = "hov.SHHMB.rmcb24.merged")
```

```{r eval=FALSE, echo=FALSE}
# scale data step is not necessary as RunFastMNN does its own scaling
rmcb24.hovShhMB.merged.clust <- NormalizeData(rmcb24.hovShhMB.merged)
rmcb24.hovShhMB.merged.clust <- FindVariableFeatures(rmcb24.hovShhMB.merged.clust)
rmcb24.hovShhMB.merged.clust <- RunFastMNN(object.list = SplitObject(rmcb24.hovShhMB.merged.clust, split.by = "orig.ident"))
rmcb24.hovShhMB.merged.clust <- RunUMAP(rmcb24.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
rmcb24.hovShhMB.merged.clust <- FindNeighbors(rmcb24.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
rmcb24.hovShhMB.merged.clust <- FindClusters(rmcb24.hovShhMB.merged.clust)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(rmcb24.hovShhMB.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/RMCB24PDX.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
rmcb24.hovShhMB.merged.clust <- 
  readRDS("outputs50K/Hovestadt_outputs/RMCB24PDX.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotHOVSHHMBtumoursRMCB24PDXfastMNN, fig.height=6, fig.width=6}
p6 <- DimPlot(rmcb24.hovShhMB.merged.clust, group.by = c("orig.ident"), pt.size = 3) +
  theme(legend.text = element_text(size = 20),
        plot.title = element_blank())
p6
```


***Calculate Mixing Metric for SHH-MB patient tumours integrated with other samples***

```{r eval=FALSE, echo=FALSE}
# calculate mixing metric for 3 SHH-MB tumours with DAOY-SPHEROID cells
# MixingMetric is a numeric vector
max.k <- 300
dySpher.hovShhMB.merged.clust.mm <- max.k - MixingMetric(object = dySpher.hovShhMB.merged.clust, 
             grouping.var = "orig.ident",
             reduction = "umap",
             dims = 1:2,
             k = 5,
             max.k = max.k,
             eps = 0,
             verbose = TRUE)
```

```{r eval=FALSE, echo=FALSE}
# calculate mixing metric for 3 SHH-MB tumours with DAOY-only cells in DAOY-CBO coculture
# MixingMetric is a numeric vector
max.k <- 300
hov.3shhMbtumours.dyOnly.merged.clust.mm <- max.k - MixingMetric(object = hov.3shhMbtumours.dyOnly.merged.clust, 
             grouping.var = "orig.ident",
             reduction = "umap",
             dims = 1:2,
             k = 5,
             max.k = max.k,
             eps = 0,
             verbose = TRUE)
```

```{r eval=FALSE, echo=FALSE}
# calculate mixing metric for 3 SHH-MB patient tumours only
# MixingMetric is a numeric vector
max.k <- 300
hovShhMB.merged.clust.mm <- max.k - MixingMetric(object = hovShhMB.merged.clust, 
             grouping.var = "orig.ident",
             reduction = "umap",
             dims = 1:2,
             k = 5,
             max.k = max.k,
             eps = 0,
             verbose = TRUE)
```

```{r eval=FALSE, echo=FALSE}
# calculate mixing metric for 3 SHH-MB tumours with RMCB18-PDX
# MixingMetric is a numeric vector
max.k <- 300
rmcb18.hovShhMB.merged.clust.mm <- max.k - MixingMetric(object = rmcb18.hovShhMB.merged.clust, 
             grouping.var = "orig.ident",
             reduction = "umap",
             dims = 1:2,
             k = 5,
             max.k = max.k,
             eps = 0,
             verbose = TRUE)
```

```{r eval=FALSE, echo=FALSE}
# calculate mixing metric for 3 SHH-MB tumours with RMCB24-PDX
# MixingMetric is a numeric vector
max.k <- 300
rmcb24.hovShhMB.merged.clust.mm <- max.k - MixingMetric(object = rmcb24.hovShhMB.merged.clust, 
             grouping.var = "orig.ident",
             reduction = "umap",
             dims = 1:2,
             k = 5,
             max.k = max.k,
             eps = 0,
             verbose = TRUE)
```

```{r eval=FALSE, echo=FALSE}
# check the lengths of each vector to find the longest - longest vector has  values
length(dySpher.hovShhMB.merged.clust.mm)   # spher + patient MB
length(hov.3shhMbtumours.dyOnly.merged.clust.mm) # cocult-only + patient MB
length(hovShhMB.merged.clust.mm) # patient MB only
length(rmcb18.hovShhMB.merged.clust.mm) # RMCB18 (PDX1) + patient MB
length(rmcb24.hovShhMB.merged.clust.mm) # RMCB24 (PDX2) + patient MB
```

```{r eval=FALSE, echo=FALSE}
hov.3shhMbtumours.dyOnly.merged.clust.mm.na <- c(hov.3shhMbtumours.dyOnly.merged.clust.mm, rep(NA, 378))
hovShhMB.merged.clust.mm.na <- c(hovShhMB.merged.clust.mm, rep(NA, 649))
rmcb18.hovShhMB.merged.clust.mm.na <- c(rmcb18.hovShhMB.merged.clust.mm, rep(NA, 483))
rmcb24.hovShhMB.merged.clust.mm.na <- c(rmcb24.hovShhMB.merged.clust.mm, rep(NA, 512))
```


```{r eval=FALSE, echo=FALSE}
summary(dySpher.hovShhMB.merged.clust.mm) # spher + patient MB
summary(hov.3shhMbtumours.dyOnly.merged.clust.mm.na) # cocult + patient MB
summary(hovShhMB.merged.clust.mm.na) # patient MB only
summary(rmcb18.hovShhMB.merged.clust.mm.na) # PDX1 + patient MB
summary(rmcb24.hovShhMB.merged.clust.mm.na) # PDX2 + patient MB
```

```{r eval=FALSE, echo=FALSE}
# create a dataframe of the MixingMetric scores for all integrations
daoy.pdx.patientShhMB.df <- data.frame(DAOY_coculture_patientMB = hov.3shhMbtumours.dyOnly.merged.clust.mm.na,
                                   DAOY_spheroid_patientMB = dySpher.hovShhMB.merged.clust.mm,
                                   patientMB = hovShhMB.merged.clust.mm.na,
                                   PDX1_patientMB = rmcb18.hovShhMB.merged.clust.mm.na,
                                   PDX2_patientMB = rmcb24.hovShhMB.merged.clust.mm.na)

daoy.pdx.patientShhMB.df.long <- pivot_longer(daoy.pdx.patientShhMB.df,
                                       cols = c(DAOY_coculture_patientMB, 
                                                DAOY_spheroid_patientMB, 
                                                patientMB,
                                                PDX1_patientMB,
                                                PDX2_patientMB),
                                       names_to = "groups.fct",
                                       values_to = "mixing_metric")

levels.order <- c("patientMB", "PDX1_patientMB", "PDX2_patientMB", "DAOY_coculture_patientMB", "DAOY_spheroid_patientMB")

daoy.pdx.patientShhMB.df.long$groups.fct <- factor(daoy.pdx.patientShhMB.df.long$groups.fct,
                                            levels = levels.order)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(daoy.pdx.patientShhMB.df.long,
        file = "./outputs50K/Hovestadt_outputs/LongDataFrame.MixingMetrics.HOVpatientSHHMB.PDX.DAOYonlyCocult.DAOYspher.rds")
```

```{r eval=FALSE, echo=FALSE}
# load the df in long format for ggplot plotting
daoy.pdx.patientShhMB.df.long <- 
  readRDS("outputs50K/Hovestadt_outputs/LongDataFrame.MixingMetrics.HOVpatientSHHMB.PDX.DAOYonlyCocult.DAOYspher.rds")
```

```{r eval=FALSE, echo=FALSE, ViolinplotMixingMetricPatientShhMBpdxDAOY, fig.height=6, fig.width=6}
# violin plot of the MixingMetric scores for all integrations
ggplot(daoy.pdx.patientShhMB.df.long, aes(x = groups.fct, y = mixing_metric)) +
  geom_violin(scale = "width", fill = "light blue", colour = "light blue") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 20),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 20, margin = margin(r=20)),
        axis.title.x = element_blank() )
```



------------------------------------------------------------------------------------------------------
***Integrate SHH-MB, PDX1, PDX2, ONS76-SPHEROID, ONS76-only coculture using fastMNN***

****Prepare the ONS76-only coculture cells for merge and fastMNN integration****
```{r eval=FALSE, echo=FALSE}
# create Seurat objects for tumours
o76only.o76cbo <- readRDS("outputs50K/ONS76-CBO/ONS76OnlyCells.SubsetONS76cboSample.RiboRetained.NoMito.SCT.CCdiffRegressed.rds")
```

```{r eval=FALSE, echo=FALSE}
head(WhichCells(o76only.o76cbo))
```

```{r eval=FALSE, echo=FALSE}
o76cbo <- readRDS("outputs50K/ONS76-CBO/filt.MitoCellsGenes.RiboGenesRetained.seurat.ONS76cbo.rds")
```

```{r eval=FALSE, echo=FALSE}
# get the DAOY only cells that have NOT been through SCT normalisation
o76onlyIDs <- WhichCells(o76only.o76cbo)
o76only <- subset(o76cbo, cells = o76onlyIDs)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(o76only,
        "./outputs50K/INTEGRATE/ONS76only.ONS76cbo.NOTclustered.NoSCT.FiltMitoCellsGenes.RiboRetained.rds")
```

```{r eval=FALSE, echo=FALSE}
# load ONS76-only coculture object
o76only <- 
  readRDS("outputs50K/INTEGRATE/ONS76only.ONS76cbo.NOTclustered.NoSCT.FiltMitoCellsGenes.RiboRetained.rds")
```

----------------------------------------------------------------------------------------------------------

```{r eval=FALSE, echo=FALSE}
# load ONS76-spheroid object
o76spher <- readRDS("outputs50K/ONS76_SPHER/filt.MitoCellsGenes.RiboGenesRetained.seurat.ONS76spher.rds")
```


***Run fastMNN on ONS76-spheroid or ONS76-only coculture in comparison to SHH-MB's only before calculating Mixing Metric***

```{r eval=FALSE, echo=FALSE}
# merge ONS76-spher with patient SHH-MB's only
o76spher.hovShhMB.merged <- merge(o76spher,
                                 y = c(muv41.shh, sj454.shh, sj577.shh),
                                 project = "hov.SHHMB.o76spher.merged")
```

```{r eval=FALSE, echo=FALSE}
metadata <- o76spher.hovShhMB.merged@meta.data
# change orig.ident as below
metadata$orig.ident[which(str_detect(metadata$orig.ident, "ons76spher"))] <- "ONS76_spheroid"
```

```{r eval=FALSE, echo=FALSE}
o76spher.hovShhMB.merged@meta.data <- metadata
```

```{r eval=FALSE, echo=FALSE}
# scale data step is not necessary as RunFastMNN does its own scaling
o76spher.hovShhMB.merged.clust <- NormalizeData(o76spher.hovShhMB.merged)
o76spher.hovShhMB.merged.clust <- FindVariableFeatures(o76spher.hovShhMB.merged.clust)
o76spher.hovShhMB.merged.clust <- RunFastMNN(object.list = SplitObject(o76spher.hovShhMB.merged.clust, split.by = "orig.ident"))
o76spher.hovShhMB.merged.clust <- RunUMAP(o76spher.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
o76spher.hovShhMB.merged.clust <- FindNeighbors(o76spher.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
o76spher.hovShhMB.merged.clust <- FindClusters(o76spher.hovShhMB.merged.clust)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(o76spher.hovShhMB.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/ONS76spheroid.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
o76spher.hovShhMB.merged.clust <- 
  readRDS("outputs50K/Hovestadt_outputs/ONS76spheroid.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotHOVSHHMBtumoursONS76spherMNN, fig.height=6, fig.width=6}
p7 <- DimPlot(o76spher.hovShhMB.merged.clust, group.by = c("orig.ident"), pt.size = 3) +
  theme(legend.text = element_text(size = 20),
        plot.title = element_blank())
p7
```

```{r eval=FALSE, echo=FALSE}
# merge ONS76-only coculture with patient SHH-MB
o76only.hovShhMB.merged <- merge(o76only,
                                 y = c(muv41.shh, sj454.shh, sj577.shh),
                                 project = "hov.SHHMB.o76only.merged")
```

```{r eval=FALSE, echo=FALSE}
metadata <- o76only.hovShhMB.merged@meta.data
# change orig.ident as below
metadata$orig.ident[which(str_detect(metadata$orig.ident, "o76cbo"))] <- "ONS76_coculture"
```

```{r eval=FALSE, echo=FALSE}
o76only.hovShhMB.merged@meta.data <- metadata
```

```{r eval=FALSE, echo=FALSE}
# scale data step is not necessary as RunFastMNN does its own scaling
o76only.hovShhMB.merged.clust <- NormalizeData(o76only.hovShhMB.merged)
o76only.hovShhMB.merged.clust <- FindVariableFeatures(o76only.hovShhMB.merged.clust)
o76only.hovShhMB.merged.clust <- RunFastMNN(object.list = SplitObject(o76only.hovShhMB.merged.clust, split.by = "orig.ident"))
o76only.hovShhMB.merged.clust <- RunUMAP(o76only.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
o76only.hovShhMB.merged.clust <- FindNeighbors(o76only.hovShhMB.merged.clust, reduction = "mnn", dims = 1:20)
o76only.hovShhMB.merged.clust <- FindClusters(o76only.hovShhMB.merged.clust)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(o76only.hovShhMB.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/ONS76onlyCoculture.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
o76only.hovShhMB.merged.clust <- 
  readRDS("outputs50K/Hovestadt_outputs/ONS76onlyCoculture.SHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotHOVSHHMBtumoursONS76onlyCocultureMNN, fig.height=6, fig.width=6}
p8 <- DimPlot(o76only.hovShhMB.merged.clust, group.by = c("orig.ident"), pt.size = 3) +
  theme(legend.text = element_text(size = 20),
        plot.title = element_blank())
p8
```

***Calculate Mixing Metrics for ONS76 samples
```{r eval=FALSE, echo=FALSE}
# calculate mixing metric for 3 SHH-MB tumours with ONS76-SPHEROID cells
# MixingMetric is a numeric vector
max.k <- 300
o76spher.hovShhMB.merged.clust.mm <- max.k - MixingMetric(object = o76spher.hovShhMB.merged.clust, 
             grouping.var = "orig.ident",
             reduction = "umap",
             dims = 1:2,
             k = 5,
             max.k = max.k,
             eps = 0,
             verbose = TRUE)
```

```{r eval=FALSE, echo=FALSE}
# calculate mixing metric for 3 SHH-MB tumours with ONS76-only cells in ONS76-CBO coculture
# MixingMetric is a numeric vector
max.k <- 300
o76only.hovShhMB.merged.clust.mm <- max.k - MixingMetric(object = o76only.hovShhMB.merged.clust, 
             grouping.var = "orig.ident",
             reduction = "umap",
             dims = 1:2,
             k = 5,
             max.k = max.k,
             eps = 0,
             verbose = TRUE)
```
START HERE - GET LENGTHS OF VECTORS
```{r eval=FALSE, echo=FALSE}
# check the lengths of each vector to find the longest - longest vector has  values
length(o76spher.hovShhMB.merged.clust.mm)   # spher + patient MB
length(o76only.hovShhMB.merged.clust.mm) # cocult-only + patient MB
length(hovShhMB.merged.clust.mm) # patient MB only
length(rmcb18.hovShhMB.merged.clust.mm) # RMCB18 (PDX1) + patient MB
length(rmcb24.hovShhMB.merged.clust.mm) # RMCB24 (PDX2) + patient MB
```

```{r eval=FALSE, echo=FALSE}
o76only.hovShhMB.merged.clust.mm.na <- c(o76only.hovShhMB.merged.clust.mm, rep(NA, 815))
hovShhMB.merged.clust.mm.na <- c(hovShhMB.merged.clust.mm, rep(NA, 1164))
rmcb18.hovShhMB.merged.clust.mm.na <- c(rmcb18.hovShhMB.merged.clust.mm, rep(NA, 998))
rmcb24.hovShhMB.merged.clust.mm.na <- c(rmcb24.hovShhMB.merged.clust.mm, rep(NA, 1027))
```

```{r eval=FALSE, echo=FALSE}
summary(o76spher.hovShhMB.merged.clust.mm) # spher + patient MB
summary(o76only.hovShhMB.merged.clust.mm.na) # cocult + patient MB
summary(hovShhMB.merged.clust.mm.na) # patient MB only
summary(rmcb18.hovShhMB.merged.clust.mm.na) # PDX1 + patient MB
summary(rmcb24.hovShhMB.merged.clust.mm.na) # PDX2 + patient MB
```

```{r eval=FALSE, echo=FALSE}
# create a dataframe of the MixingMetric scores for all integrations
o76.pdx.patientShhMB.df <- data.frame(ONS76_coculture_patientMB = o76only.hovShhMB.merged.clust.mm.na,
                                   ONS76_spheroid_patientMB = o76spher.hovShhMB.merged.clust.mm,
                                   patientMB = hovShhMB.merged.clust.mm.na,
                                   PDX1_patientMB = rmcb18.hovShhMB.merged.clust.mm.na,
                                   PDX2_patientMB = rmcb24.hovShhMB.merged.clust.mm.na)

o76.pdx.patientShhMB.df.long <- pivot_longer(o76.pdx.patientShhMB.df,
                                       cols = c(ONS76_coculture_patientMB, 
                                                ONS76_spheroid_patientMB, 
                                                patientMB,
                                                PDX1_patientMB,
                                                PDX2_patientMB),
                                       names_to = "groups.fct",
                                       values_to = "mixing_metric")

levels.order <- c("patientMB", "PDX1_patientMB", "PDX2_patientMB", "ONS76_coculture_patientMB", "ONS76_spheroid_patientMB")

o76.pdx.patientShhMB.df.long$groups.fct <- factor(o76.pdx.patientShhMB.df.long$groups.fct,
                                            levels = levels.order)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(o76.pdx.patientShhMB.df.long,
        file = "./outputs50K/Hovestadt_outputs/LongDataFrame.MixingMetrics.HOVpatientSHHMB.PDX.ONS76onlyCocult.ONS76spher.rds")
```

```{r eval=FALSE, echo=FALSE}
# load the df in long format for ggplot plotting
o76.pdx.patientShhMB.df.long <- 
  readRDS("outputs50K/Hovestadt_outputs/LongDataFrame.MixingMetrics.HOVpatientSHHMB.PDX.ONS76onlyCocult.ONS76spher.rds")
```

```{r eval=FALSE, echo=FALSE, ViolinplotMixingMetricPatientShhMBpdxONS76, fig.height=6, fig.width=6}
# violin plot of the MixingMetric scores for all integrations
ggplot(o76.pdx.patientShhMB.df.long, aes(x = groups.fct, y = mixing_metric)) +
  geom_violin(scale = "width", fill = "light blue", colour = "light blue") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 20),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 20, margin = margin(r=20)),
        axis.title.x = element_blank() )
```

****Create one dataframe for both cell lines****
```{r eval=FALSE, echo=FALSE}
# check the lengths of each vector to find the longest - longest vector has  values
length(dySpher.hovShhMB.merged.clust.mm)   # spher + patient MB 
length(hov.3shhMbtumours.dyOnly.merged.clust.mm) # cocult-only + patient MB
length(hovShhMB.merged.clust.mm) # patient MB only
length(rmcb18.hovShhMB.merged.clust.mm) # RMCB18 (PDX1) + patient MB
length(rmcb24.hovShhMB.merged.clust.mm) # RMCB24 (PDX2) + patient MB
length(o76spher.hovShhMB.merged.clust.mm)   # spher + patient MB
length(o76only.hovShhMB.merged.clust.mm) # cocult-only + patient MB
```

```{r eval=FALSE, echo=FALSE}
hov.3shhMbtumours.dyOnly.merged.clust.mm.na <- c(hov.3shhMbtumours.dyOnly.merged.clust.mm, rep(NA, 893))
dySpher.hovShhMB.merged.clust.mm.na <- c(dySpher.hovShhMB.merged.clust.mm, rep(NA, 515))
hovShhMB.merged.clust.mm.na <- c(hovShhMB.merged.clust.mm, rep(NA, 1164))
rmcb18.hovShhMB.merged.clust.mm.na <- c(rmcb18.hovShhMB.merged.clust.mm, rep(NA, 998))
rmcb24.hovShhMB.merged.clust.mm.na <- c(rmcb24.hovShhMB.merged.clust.mm, rep(NA, 1027))
o76only.hovShhMB.merged.clust.mm.na <- c(o76only.hovShhMB.merged.clust.mm, rep(NA, 815))
```

```{r eval=FALSE, echo=FALSE}
summary(hovShhMB.merged.clust.mm.na) # patient MB only
summary(rmcb18.hovShhMB.merged.clust.mm.na) # PDX1 + patient MB
summary(rmcb24.hovShhMB.merged.clust.mm.na) # PDX2 + patient MB
summary(hov.3shhMbtumours.dyOnly.merged.clust.mm.na) # DAOY-cocult + patient MB
summary(o76only.hovShhMB.merged.clust.mm.na) # ONS76-cocult + patient MB
summary(dySpher.hovShhMB.merged.clust.mm.na) # DAOY-spher + patient MB
summary(o76spher.hovShhMB.merged.clust.mm) # ONS76-spher + patient MB
```

```{r eval=FALSE, echo=FALSE}
mixmetSumm <- rbind(summary(hovShhMB.merged.clust.mm), 
      summary(rmcb18.hovShhMB.merged.clust.mm),
      summary(rmcb24.hovShhMB.merged.clust.mm),
      summary(hov.3shhMbtumours.dyOnly.merged.clust.mm),
      summary(o76only.hovShhMB.merged.clust.mm),
      summary(dySpher.hovShhMB.merged.clust.mm),
      summary(o76spher.hovShhMB.merged.clust.mm))
mixmetSumm
```

```{r eval=FALSE, echo=FALSE}
colnames(mixmetSumm)
```

```{r eval=FALSE, echo=FALSE}
mixmetSumm.df <- as.data.frame(mixmetSumm)
```

```{r eval=FALSE, echo=FALSE}
head(mixmetSumm.df)
```


```{r eval=FALSE, echo=FALSE}
# add a 'sample' column to the above df at the beginning
Sample <- c("patientMB", "PDX1_patientMB", "PDX2_patientMB", "DAOY_coculture_patientMB",  "ONS76_coculture_patientMB", "DAOY_spheroid_patientMB", "ONS76_spheroid_patientMB")

mixmetSumm.df2 <- cbind(Sample, mixmetSumm.df)

mixmetSumm.df2
```

```{r eval=FALSE, echo=FALSE}
write.csv(mixmetSumm.df2,
          file = "./outputs50K/Hovestadt_outputs/MixingMetricTableOfSummaryFunctionValues.7conditions.csv",
          quote = FALSE)
```


```{r eval=FALSE, echo=FALSE}
# create a dataframe of the MixingMetric scores for all integrations
daoy.o76.pdx.patientShhMB.df <- data.frame(DAOY_coculture_patientMB = hov.3shhMbtumours.dyOnly.merged.clust.mm.na,
                                   DAOY_spheroid_patientMB = dySpher.hovShhMB.merged.clust.mm.na,
                                   patientMB = hovShhMB.merged.clust.mm.na,
                                   PDX1_patientMB = rmcb18.hovShhMB.merged.clust.mm.na,
                                   PDX2_patientMB = rmcb24.hovShhMB.merged.clust.mm.na,
                                   ONS76_coculture_patientMB = o76only.hovShhMB.merged.clust.mm.na,
                                   ONS76_spheroid_patientMB = o76spher.hovShhMB.merged.clust.mm)

daoy.o76.pdx.patientShhMB.df.long <- pivot_longer(daoy.o76.pdx.patientShhMB.df,
                                       cols = c(DAOY_coculture_patientMB, 
                                                DAOY_spheroid_patientMB, 
                                                patientMB,
                                                PDX1_patientMB,
                                                PDX2_patientMB,
                                                ONS76_coculture_patientMB,
                                                ONS76_spheroid_patientMB),
                                       names_to = "groups.fct",
                                       values_to = "mixing_metric")

levels.order <- c("patientMB", "PDX1_patientMB", "PDX2_patientMB", "DAOY_coculture_patientMB",  "ONS76_coculture_patientMB", "DAOY_spheroid_patientMB", "ONS76_spheroid_patientMB")

daoy.o76.pdx.patientShhMB.df.long$groups.fct <- factor(daoy.o76.pdx.patientShhMB.df.long$groups.fct,
                                            levels = levels.order)
```

```{r eval=FALSE, echo=FALSE}
# save the df in long format for ggplot plotting
saveRDS(daoy.o76.pdx.patientShhMB.df.long,
  file = "./outputs50K/Hovestadt_outputs/LongDataFrame.MixingMetrics.HOVpatientSHHMB.PDX.DAOYonlyCocult.DAOYspher.ONS76onlyCocult.ONS76spher.rds")
```

```{r eval=FALSE, echo=FALSE}
# load the long df
daoy.o76.pdx.patientShhMB.df.long <- 
  readRDS("outputs50K/Hovestadt_outputs/LongDataFrame.MixingMetrics.HOVpatientSHHMB.PDX.DAOYonlyCocult.DAOYspher.ONS76onlyCocult.ONS76spher.rds")
```

```{r eval=FALSE, echo=FALSE, ViolinplotMixingMetricPatientShhMBpdxONS76andDAOY, fig.height=6, fig.width=12}
# violin plot of the MixingMetric scores for all integrations
ggplot(daoy.o76.pdx.patientShhMB.df.long, aes(x = groups.fct, y = mixing_metric)) +
  geom_violin(scale = "width", fill = "light blue", colour = "light blue") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 20),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 20, margin = margin(r=20)),
        axis.title.x = element_blank() )
```




########################################################################################################
--------------------------------------------------------------------------------------------------------
CODE THAT WAS NOT VERY USEFUL BUT RETAINED FOR FUTURE USE


```{r eval=FALSE, echo=FALSE}
saveRDS(hov.6tumours.o76spher.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/ONS76spher.WNTMBSHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
hov.6tumours.o76spher.merged.clust <- 
  readRDS("outputs50K/Hovestadt_outputs/ONS76spher.WNTMBSHHMBHov.fastMNN.clust.rds")
```


```{r eval=FALSE, echo=FALSE}
# this file was prepared previously, not planning to use it!
saveRDS(hov.6tumours.dySpher.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/DAOYspher.WNTMBSHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
hov.6tumours.dySpher.merged.clust <- readRDS("outputs50K/Hovestadt_outputs/DAOYspher.WNTMBSHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
saveRDS(hov.6tumours.dyOnly.merged.clust,
        file = "./outputs50K/Hovestadt_outputs/DAOYonlyDAOYcbo.WNTMBSHHMBHov.fastMNN.clust.rds")
```

```{r eval=FALSE, echo=FALSE}
hov.6tumours.dyOnly.merged.clust <- readRDS("outputs50K/Hovestadt_outputs/DAOYonlyDAOYcbo.WNTMBSHHMBHov.fastMNN.clust.rds")
```


SILHOUETTE ANALYSIS:
INTEGRATED ANALYSIS OF PATIENT SHH-MB SAMPLES WITH DAOY SPHEROID USING SILHOUETTE ANALYSIS

```
{r}
# load the files
shhMb1.combined <- readRDS("outputs50K/Hovestadt_outputs/object_shhMb1.combined.Hovestadt.SHH-MBpatients_n=3.DAOYmono.spher.cocult.clustered.conventionalIntegration.rds")
shhMb1.1.combined <- readRDS("outputs50K/Hovestadt_outputs/object_shhMb1.1.combined.Hovestadt.SHH-MBpatients_n=3.DAOYspher.cocult.clustered.conventionalIntegration.rds")
shhMb1.2.combined <- readRDS("outputs50K/Hovestadt_outputs/object_shhMb1.2.combined.Hovestadt.SHH-MBpatients_n=3.DAOYcocult.clustered.conventionalIntegration.rds")
shhMb1.3.combined <- readRDS("outputs50K/Hovestadt_outputs/object_shhMb1.3.combined.Hovestadt.SHH-MBpatients_n=3.DAOYspher.clustered.conventionalIntegration.rds")
shhMb1.4.combined <- readRDS("outputs50K/Hovestadt_outputs/object_shhMb1.4.combined.Hovestadt.SHH-MBpatients_n=3.DAOYmono.clustered.conventionalIntegration.rds")
shhMb1.5.combined <- readRDS("outputs50K/Hovestadt_outputs/object_shhMb1.5.combined.Hovestadt.SHH-MBpatients_n=3.NoDAOY.conventionalIntegration.rds")
```

```
{r}
DimPlot(shhMb1.4.combined, reduction = "tsne", group.by = "seurat_clusters", pt.size = 2) +
 # guides(color = guide_legend(override.aes = list(size=12), ncol=1) ) - this is optional
  theme(legend.text=element_text(size=20)) +
  labs(title = "")
```

```
{r}
# Calculate Euclidean distance matrix -  GetAssayData() is used to extract the data matrix from the Seurat object, and t() is used to transpose it so that the cells are in rows and the genes are in columns, which is the format required for the dist() function.
dist.mat <- dist(t(GetAssayData(object = shhMb1.4.combined, slot = "data")))
```

```
{r}
# save the dist.mat as it takes a long time to compute - THIS FILE HAS BEEN DELETED
saveRDS(dist.mat,
        file = "./outputs50K/Hovestadt_outputs/distanceMatrix.shhMb1.4.combined.Hovestadt.SHH-MBpatients_n=3.DAOYmono.clustered.conventionalIntegration.rds")
```

```
{r}
dist.mat <- readRDS("outputs50K/Hovestadt_outputs/distanceMatrix.shhMb1.4.combined.Hovestadt.SHH-MBpatients_n=3.DAOYmono.clustered.conventionalIntegration.rds")
clusters <- shhMb1.4.combined@meta.data$integrated_snn_res.0.5
```

```
{r}
# Calculate silhouette scores
sil <- silhouette(x = as.numeric(x = as.factor(x = clusters)), dist = dist.mat)
```

```
{r}
shhMb1.4.combined$sil <- sil[, 3]
```

```
{r}
# median scores that are lower indicate better cell mixing, i.e. cells are more similar
summary(shhMb1.4.combined@meta.data$sil)
```




SCTransform integration did not work so well even with cell cycle regression and using the full DAOY-monolayer (DAOY-rep) or using a donsampled DAOY monolayer 

Try SCTransform as the spheroid and the daoy-only coculture overlap a lot

```
{r}
# trial of SCTransform
dyMono <- readRDS("outputs50K/DAOY_2D/filt.RiboMitoCellsGenes.seurat.DaoyMonolayer.rds")
# normalise data
dyMono <- NormalizeData(dyMono)
# calculate CC difference
dyMono <- CellCycleScoring(dyMono,
                g2m.features = cc.genes$g2m.genes,
                s.features = cc.genes$s.genes)
dyMono$CC.difference <- dyMono$S.Score - dyMono$G2M.Score
```

```
{r}

# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
dyMono.input.integration <- SCTransform(dyMono, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```
{r}
dySpher.input.integration <- 
  readRDS("outputs50K/DAOY_SPHER/DAOYspher.inputIntegration.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```



As can be seen from the below, the dyOnly.dyCbo object has a prefix before the cell ID. This must be removed before proceeding
```
{r}
head(WhichCells(dyOnly.dyCbo))
```

```
{r}
# store the above cell IDs as a vector
dyOnly.cellID <- WhichCells(dyOnly.dyCbo)
```

```
{r}
# remove the prefixes from the vector cell names and save as a new object of cell IDs
dyOnly.cellID.corrected <- gsub('dyCbo.data-', '', dyOnly.cellID)
head(dyOnly.cellID.corrected) # the prefixes have been removed
```

```
{r}
# rename the cell IDs with the corrected cell names (.ccn = corrected cell names)
# the object dyOnly.dyCbo.ccn is the one to use going forwards
dyOnly.dyCbo.ccn <- RenameCells(dyOnly.dyCbo, new.names = dyOnly.cellID.corrected)
head(WhichCells(dyOnly.dyCbo.ccn))
```

```
{r}
# prepare MUV41
# normalise the object
muv41 <- NormalizeData(muv41)

# get cell cycle scores, then regress this variable out
muv41 <- CellCycleScoring(muv41,
                          s.features = cc.genes$s.genes,
                          g2m.features = cc.genes$g2m.genes)
muv41$CC.difference <- muv41$S.Score - muv41$G2M.Score
```

```
{r}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
muv41.input.integration <- SCTransform(muv41, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```
{r}
# prepare SJ454
# normalise the object
sj454 <- NormalizeData(sj454)

# get cell cycle scores, then regress this variable out
sj454 <- CellCycleScoring(sj454,
                          s.features = cc.genes$s.genes,
                          g2m.features = cc.genes$g2m.genes,
                          set.ident = TRUE)
sj454$CC.difference <- sj454$S.Score - sj454$G2M.Score
```

```
{r}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
sj454.input.integration <- SCTransform(sj454, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```
{r}
# prepare SJ577
# normalise the object
sj577 <- NormalizeData(sj577)

# get cell cycle scores, then regress this variable out
sj577 <- CellCycleScoring(sj577,
                          s.features = cc.genes$s.genes,
                          g2m.features = cc.genes$g2m.genes,
                          set.ident = TRUE)
sj577$CC.difference <- sj577$S.Score - sj577$G2M.Score
```

```
{r}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
sj577.input.integration <- SCTransform(sj577, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```
{r}
# do the SCT integration 
dy.sct.list <- list(dyMono.input.integration, 
                dySpher.input.integration,
                dyOnly.dyCbo.ccn,
                muv41.input.integration, 
                sj454.input.integration,
                sj577.input.integration)
features <- SelectIntegrationFeatures(object.list = dy.sct.list, nfeatures = 3000)
dy.sct.list <- PrepSCTIntegration(object.list = dy.sct.list, anchor.features = features)
```

```
{r}
dy.sct.anchors <- FindIntegrationAnchors(object.list = dy.sct.list, normalization.method = "SCT",
    anchor.features = features)
dy.combined.sct <- IntegrateData(anchorset = dy.sct.anchors, normalization.method = "SCT")
```

```
{r}
dy.combined.sct <- RunPCA(dy.combined.sct, verbose = FALSE)
dy.combined.sct <- RunTSNE(dy.combined.sct, reduction = "pca", dims = 1:30)
```

```
{r}
# conventional integration workflow is better!
DimPlot(dy.combined.sct, reduction = "tsne", group.by = "orig.ident")
```









