---
title: "ONS76-SPHER_Seurat_EXPLORE1_50K"
author: "JJ"
date: "19/01/2023"
output: html_document
---

```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE,
                      fig.path = "./outputs50K/ONS76_SPHER/figures/")
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(glmGamPoi)
library(presto)
library(clustree)
library(tidyverse)
library(magrittr)
library(patchwork)
})
```


```{r}
# load the data from scratch
# load the ONS76-2D raw data (from "working_data" folder)
data_dir <- "./working_data/ONS76-SPHER/"   # the dot is current dir and is essential!
list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r}
# note these are .gz files which can be directly read. n = 3838 cells
o76spher.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                         features = paste0(data_dir, "features.tsv.gz"), 
                         cells = paste0(data_dir, "barcodes.tsv.gz"))
# create the seurat object, keep genes expressed in 10 or more cells
o76spher <- CreateSeuratObject(counts = o76spher.data,
                          min.cells = 10, min.genes=200,
                          project = "ons76spher")

dim(o76spher) # 17,677 genes and 1322 cells
```

```{r}
# identify mitochondrial genes
mito.genes <- grep(pattern = "^MT", x = rownames(x = o76spher), value = TRUE)
percent.mito <- 
  Matrix::colSums(o76spher[mito.genes, ])/Matrix::colSums(o76spher)

# identify ribosomal genes
ribo.genes <- grep(pattern = "^RP[SL]", x = rownames(x = o76spher), value = TRUE)
percent.ribo <- 
  Matrix::colSums(o76spher[ribo.genes, ])/Matrix::colSums(o76spher)

# add the ribosome and mitochondrial gene percent to the metadata
o76spher <- 
  AddMetaData(object = o76spher, metadata = percent.mito, col.name = "percent.mito")

o76spher <- 
  AddMetaData(object = o76spher, metadata = percent.ribo, col.name = "percent.ribo")

# remove ribosomal and mitochondrial genes
# remove ribosomal genes first
o76spher <- o76spher[ ! grepl('^RP[SL]', rownames(o76spher)), ]

# remove mitochondrial genes 
o76spher <- o76spher[ ! grepl("^MT-", rownames(o76spher)), ]

dim(o76spher) # 17,560 genes and 1322 cells
```

```{r}
# remove all cells with percent.mito > 10%
o76spher <- subset(o76spher, subset = percent.mito < 0.1)
dim(o76spher) # reduced to 17,569 genes and 1164 cells
```

```{r}
# save the NON-normalised, NON-clustered cbo seurat object
saveRDS(o76spher, file = "./outputs50K/ONS76_SPHER/filt.RiboMitoCellsGenes.seurat.ons76spheroid.rds")
```

```{r}
o76spher <- readRDS("outputs50K/ONS76_SPHER/filt.RiboMitoCellsGenes.seurat.ons76spheroid.rds")
```

```{r}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference

# normalise object before running cell cycle scores as per Seurat github issue: https://github.com/satijalab/seurat/issues/1679
o76spher <- NormalizeData(o76spher)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
o76spher <- CellCycleScoring(o76spher,
                g2m.features = g2m.genes,
                s.features = s.genes)
o76spher$CC.difference <- o76spher$S.Score - o76spher$G2M.Score
```

```{r}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
o76spher.input.integration <- SCTransform(o76spher, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```{r}
# save the SCTransformed object, CC difference regressed, PCA reduction, as input for integrated workflow
saveRDS(o76spher.input.integration, "./outputs50K/ONS76_SPHER/ONS76spher.inputIntegration.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```

```{r}
o76spher <- SCTransform(o76spher, vars.to.regress = "CC.difference", method = "glmGamPoi", 
                        verbose = FALSE) %>%   
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.5)
```

```{r}
# save the clustered, SCTransformed object above
saveRDS(o76spher, 
        file = "./outputs50K/ONS76_SPHER/ons76spher.RiboMitoCellsGenesFilt.SCT.clustered.CCdiffRegressed.rds")
```

```{r}
# load the above object. This object not yet been annotated.
o76spher <- 
  readRDS("outputs50K/ONS76_SPHER/ons76spher.RiboMitoCellsGenesFilt.SCT.clustered.CCdiffRegressed.rds")
```

```{r DimplotCboUmapSCTransform, fig.height=6, fig.width=10}
# for res = 0.5, 5 clusters are found
DimPlot(o76spher, label = TRUE)
```

```{r}
# get the markers of each cluster
o76spher.POSmarkers <- FindAllMarkers(object = o76spher, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
o76spher.POSmarkers %>% 
  dplyr::group_by(cluster) %>% 
  slice_head(n = 3)
```

```{r}
# MEIS1 and CNTN1 are expressed of these five
o76spher.POSmarkers %>%
  filter(gene %in% c("NEUROD1", "CNTN1", "CNTN2", "MEIS1", "DCX1"))
```

```{r}
saveRDS(o76spher.POSmarkers,
        file = "./outputs50K/ONS76_SPHER/POS.markers.ONS76spheroid.SCT.clustered.rds")
```

```{r}
o76spher.POSmarkers <- readRDS("outputs50K/ONS76_SPHER/POS.markers.ONS76spheroid.SCT.clustered.rds")
```

```{r HeatmapONS76spherTop10markers5clustersRes0.5, fig.height=6, fig.width=6}
# first load o76spher seurat object above and the file of positive markers, then only run this chunk
o76spher.POSmarkers %>% 
  group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(o76spher, features = top10$gene) + NoLegend()
```

As can be seen in the above heatmap, the markers do not do a great job of separating the 5 clusters, e.g. clusters 2 and 3 have quite similar gene expression. Also looking at clustree output, at res=0.4 one of the clusters looks unstable. Try res=0.4 nevertheless.
```{r}
# resolution = 0.4
# first load "filt.RiboMitoCellsGenes.seurat.ons76spheroid.rds" then normalise, run CC regression,
# then cluster the cells as per this code chunk
o76spher.res0.4 <- SCTransform(o76spher, vars.to.regress = "CC.difference", method = "glmGamPoi", verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.4)
```

```{r}
# get the markers of each cluster 
o76spher.POSmarkers.res0.4 <- 
  FindAllMarkers(object = o76spher.res0.4, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r HeatmapONS76spherTop10markers5clustersRes0.4, fig.height=6, fig.width=6}
# first load dyMono seurat object above, then only run this chunk
o76spher.POSmarkers.res0.4 %>% 
  group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(o76spher.res0.4, features = top10$gene) + guides(color = "none")
```

The heatmap at res=0.4 shows that clusters are still not that well defined. Try resolution = 0.2
```{r}
# resolution = 0.2
# first load "filt.RiboMitoCellsGenes.seurat.ons76spheroid.rds" then normalise, run CC regression,
# then cluster the cells as per this code chunk
o76spher.res0.2 <- SCTransform(o76spher, vars.to.regress = "CC.difference", method = "glmGamPoi", verbose = FALSE) %>% 
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.2)
```

```{r}
# get the markers of each cluster 
o76spher.POSmarkers.res0.2 <- 
  FindAllMarkers(object = o76spher.res0.2, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r HeatmapONS76spherTop10markers2clustersRes0.2, fig.height=6, fig.width=6}
# first load o76spher seurat object above, then only run this chunk
o76spher.POSmarkers.res0.2 %>% 
  group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(o76spher.res0.2, features = top10$gene) + guides(color = "none")
```

The above heatmap with resolution = 0.2 and 2 clusters shows the best differentiation between clusters.

```{r}
saveRDS(o76spher.res0.2,
        file = "./outputs50K/ONS76_SPHER/ons76spheroid.RiboMitoCellsGenesFilt.SCT.res0.2.clustered.CCdiffRegressed.rds")
```

```{r}
# save the POS markers at res=0.2
saveRDS(o76spher.POSmarkers.res0.2,
        file = "./outputs50K/ONS76_SPHER/POS.markers.res0.2.ONS76spheroid.SCT.clustered.rds")
```

Run presto
```{r}
# prepare the data for gProfiler (web app) input
o76spher.res0.2.genes.presto <- wilcoxauc(o76spher.res0.2, 'seurat_clusters')
nrow(o76spher.res0.2.genes.presto)
dplyr::count(o76spher.res0.2.genes.presto, group)
```

```{r}
head(o76spher.res0.2.genes.presto, 3)
```

```{r}
# filter and order the gene list for cluster 0
o76spher.res0.2.genes.presto.clust0 <- o76spher.res0.2.genes.presto %>% 
  dplyr::filter(group == "0" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.res0.2.genes.presto.clust0) # 203 genes
```

```{r}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.res0.2.genes.presto.clust0, 
          "./outputs50K/ONS76_SPHER/ONS76spheroid.res0.2.cluster0.genes.presto.gProfilerInput.csv", quote = F)
```

```{r}
# filter and order the gene list for cluster 1
o76spher.res0.2.genes.presto.clust1 <- o76spher.res0.2.genes.presto %>% 
  dplyr::filter(group == "1" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.res0.2.genes.presto.clust1) # 173 genes
```

```{r}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.res0.2.genes.presto.clust1, 
          "./outputs50K/ONS76_SPHER/ONS76spheroid.res0.2.cluster1.genes.presto.gProfilerInput.csv", quote = F)
```



***Clustree***
```{r}
# attempt to use Clustree package to find the right number of clusters
# -----------------------------------------------------------------------------------------------

# Select a range of resolutions
resolution.range <- seq(from = 0, to = 1.8, by = 0.2)

# Find clusters using a range of resolutions, the clustering information is in the metadata slot
ons76spher.clusters <- FindClusters(object = o76spher, 
                                      resolution = resolution.range,
                                      verbose = FALSE)
```

```{r ClustreeCBO, fig.height=8}
# the clustering tree finds 2-4 major clusters initially with res = 0.5.
# when res > 0.4 the cluster assignment becomes unstable.
# clusters 0-2 have pct.1 ~ pct.2, whereas clusters 3 and 4 look unique 
clustree(ons76spher.clusters)
```


```{r}
# get the number of cells per cluster
table(Idents(o76spher))
```


```{r}
# load the dataset
o76spher.res0.2 <- readRDS("outputs50K/ONS76_SPHER/ons76spheroid.RiboMitoCellsGenesFilt.SCT.res0.2.clustered.CCdiffRegressed.rds")
```

```{r}
# confirm that cell cycle phase metadata is present - it is
head(o76spher.res0.2@meta.data)
```

```{r}
# extract the G1 only cells from o76spher (this 'o76spher' object is the already CLUSTERED object)
# first, switch Idents to 'Phase'
Idents(o76spher.res0.2) <- "Phase"
o76spher.g1.cellIDs <- WhichCells(o76spher.res0.2, idents = "G1")
length(o76spher.g1.cellIDs)
```

```{r}
# these cell IDs look correctly formatted
# object o76spher has to be "filt.RiboMitoCellsGenes.seurat.ons76spheroid.rds" above!
o76spher.g1cells <- subset(o76spher, cells = o76spher.g1.cellIDs)
head(WhichCells(o76spher.g1cells))
```

```{r}
# check the number of cells is correct in the subsetted object
dim(o76spher.g1cells) # 668 cells in G1
```


```{r}
# create an object suitable for integration (don't need to regress cell cycle!)

o76spher.g1cells.input.integration <- SCTransform(o76spher.g1cells,
                                                  method = "glmGamPoi",
                                                  vst.flavor = "v2",
                                                  verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```{r}
dim(o76spher.g1cells.input.integration) # object has 668 cells which is correct
```

```{r}
# save the "o76spher.g1cells.input.integration" object
saveRDS(o76spher.g1cells.input.integration, 
        file = "./outputs50K/ONS76_SPHER/G1phaseCells.ONS76spheroid.inputIntegration.SCTransformed.PCA.unclustered.rds")
```








